# Types of Manufacturing Processes

```{r setup-01, include=FALSE}
# include the helpers.R file for pdf output of youtube on pdfs as a clickable link
source("R/helpers.R")
# include the required packages to make sure the page can be built.
source("R/required_packages.R")
```
------------------------------------------------------------------------

## Learning Objectives

By the end of this chapter, you will be able to:

1. Identify and classify the seven main types of manufacturing processes
2. Select appropriate manufacturing processes based on material, volume, and precision requirements
3. Compare cost, speed, and quality trade-offs between process types
4. Recognize manufacturing processes used in automotive, food, and defense industries
5. Apply process selection criteria to real-world manufacturing scenarios

------------------------------------------------------------------------

## Introduction to Manufacturing Processes

Manufacturing is defined as the conversion of raw materials into finished goods on a large scale using man and machine. **Manufacturing processes** are the specific methods used for this conversion. Based on product requirements, different types of manufacturing processes are selected to achieve the required output with optimal cost, quality, and efficiency.

```{r manufacturing-overview, echo=FALSE, fig.align="center", fig.cap="Overview of Manufacturing Process Categories", fig.width=10, fig.height=6}
# Create an overview visualization of manufacturing processes
process_categories <- data.frame(
  Category = c("Casting", "Molding", "Forming", "Machining", "Joining", "Coating", "Additive"),
  Primary_Material = c("Metals", "Plastics", "Metals", "All", "All", "All", "All"),
  Volume = c(4, 5, 4, 2, 3, 3, 1),
  Complexity = c(3, 4, 2, 5, 3, 2, 5),
  x_pos = c(1, 2, 3, 4, 5, 6, 7)
)

ggplot(process_categories, aes(x = x_pos, y = 1)) +
  geom_tile(aes(fill = Category), width = 0.9, height = 0.6, color = "white", linewidth = 2) +
  geom_text(aes(label = Category), color = "white", fontface = "bold", size = 4) +
  scale_fill_manual(values = c("#E74C3C", "#3498DB", "#2ECC71", "#9B59B6", "#F39C12", "#1ABC9C", "#34495E")) +
  labs(title = "Seven Main Categories of Manufacturing Processes",
       subtitle = "Each category contains multiple specific processes for different applications") +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    legend.position = "none"
  )
```

<details>
<summary>Why do we need different manufacturing processes?</summary>

Different products require different manufacturing approaches based on:

- **Material type**: Metals, plastics, ceramics, composites
- **Production volume**: One-off prototypes vs. millions of units
- **Geometric complexity**: Simple shapes vs. intricate details
- **Precision requirements**: Loose tolerances vs. micron-level accuracy
- **Cost constraints**: Budget limitations affect process selection
- **Lead time**: How quickly parts are needed

No single process is optimal for all situations!

</details>

A good video summarizing 6 of the 7 manufacturing processes is linked below (Coating/Plating is not covered):

```{r video-manufacturing-overview, echo=FALSE, results='asis'}
embed_youtube("Um_g8sQ_p3Y", "Manufacturing Processes Overview")
```

------------------------------------------------------------------------

## Process Selection Framework

Before examining each process type, it's important to understand how engineers select the right manufacturing process. The following visualization shows the key factors:

```{r process-selection-matrix, echo=FALSE, fig.align="center", fig.cap="Process Selection Based on Volume and Complexity", fig.width=10, fig.height=7}
# Create process selection visualization
selection_data <- data.frame(
  Process = c("Sand Casting", "Die Casting", "Investment Casting",
              "Injection Molding", "Blow Molding", "Thermoforming",
              "Forging", "Stamping", "Extrusion",
              "CNC Machining", "3D Printing", "Manual Assembly"),
  Volume = c(2, 5, 2, 5, 4, 3, 4, 5, 5, 2, 1, 2),
  Complexity = c(3, 3, 5, 4, 2, 2, 2, 2, 1, 5, 5, 3),
  Category = c("Casting", "Casting", "Casting",
               "Molding", "Molding", "Molding",
               "Forming", "Forming", "Forming",
               "Machining", "Additive", "Joining")
)

ggplot(selection_data, aes(x = Volume, y = Complexity, color = Category)) +
  geom_point(size = 6, alpha = 0.8) +
  geom_text(aes(label = Process), vjust = -1, size = 3, fontface = "bold") +
  scale_x_continuous(limits = c(0, 6), breaks = 1:5,
                     labels = c("Prototype\n(1-10)", "Low\n(10-1K)", "Medium\n(1K-10K)",
                               "High\n(10K-100K)", "Mass\n(100K+)")) +
  scale_y_continuous(limits = c(0, 6), breaks = 1:5,
                     labels = c("Simple", "Low", "Medium", "High", "Very High")) +
  scale_color_manual(values = c("Additive" = "#9B59B6", "Casting" = "#E74C3C", "Forming" = "#2ECC71",
                                "Joining" = "#1ABC9C", "Machining" = "#3498DB", "Molding" = "#F39C12")) +
  labs(x = "Production Volume", y = "Part Complexity",
       title = "Manufacturing Process Selection Guide",
       subtitle = "Choose processes based on volume requirements and part complexity") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r process-comparison-table, echo=FALSE}
comparison_data <- tibble(
  `Process Type` = c("Casting", "Molding", "Forming", "Machining", "Joining", "Coating", "Additive"),
  `Typical Materials` = c("Metals, alloys", "Plastics, rubber", "Metals, sheet", "All materials", "All materials", "All materials", "Plastics, metals, ceramics"),
  `Production Volume` = c("Low to High", "Medium to Very High", "Medium to Very High", "Low to Medium", "All volumes", "All volumes", "Prototype to Low"),
  `Typical Tolerance` = c("±0.5-2mm", "±0.1-0.5mm", "±0.1-1mm", "±0.01-0.1mm", "Varies", "N/A", "±0.1-0.3mm"),
  `Tooling Cost` = c("Medium-High", "High", "Medium-High", "Low", "Low", "Low", "None"),
  `Unit Cost at Volume` = c("Low", "Very Low", "Low", "High", "Medium", "Low", "High")
)

comparison_data %>%
  kable(format = "html", caption = "Manufacturing Process Comparison Summary") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered", "responsive"),
    full_width = TRUE,
    position = "center"
  ) %>%
  row_spec(0, bold = TRUE, background = "#2C3E50", color = "white") %>%
  column_spec(1, bold = TRUE, background = "#ECF0F1")
```

------------------------------------------------------------------------

## 1. Casting Processes

**Casting** is a process of pouring liquid metal into a mold containing the hollow shape of the desired outcome. It uses sprue, gates, and runners to direct the molten metal flow.

```{r casting-comparison, echo=FALSE, fig.align="center", fig.cap="Comparison of Casting Processes", fig.width=10, fig.height=5}
casting_data <- data.frame(
  Process = c("Sand Casting", "Die Casting", "Investment Casting", "Permanent Mold", "Centrifugal", "Shell Molding"),
  Tooling_Cost = c(1, 5, 3, 4, 3, 2),
  Production_Rate = c(1, 5, 2, 3, 2, 3),
  Surface_Finish = c(1, 4, 5, 3, 3, 4),
  Complexity = c(3, 3, 5, 2, 1, 3)
)

casting_long <- casting_data %>%
  pivot_longer(cols = -Process, names_to = "Metric", values_to = "Score") %>%
  mutate(Metric = gsub("_", " ", Metric))

ggplot(casting_long, aes(x = Metric, y = Score, fill = Process)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(breaks = 1:5, labels = c("Low", "Med-Low", "Medium", "Med-High", "High")) +
  labs(title = "Casting Process Comparison",
       subtitle = "Comparing key characteristics across casting methods",
       x = "", y = "Rating") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 15, hjust = 1),
    legend.position = "bottom"
  )
```

### Sand Casting

Sand casting is a metal casting process that uses sand as the mold material. This process has a low production rate as the sand mold must be destroyed to remove the part. However, it is economical for low-volume production.

**Process Steps:**

1. Create a pattern (usually wood or metal)
2. Pack sand around the pattern to form the mold
3. Remove the pattern, leaving a cavity
4. Pour molten metal into the cavity
5. Allow solidification
6. Break the sand mold to remove the casting

**Advantages:** Low tooling cost, can produce very large parts, suitable for all metals

**Disadvantages:** Poor surface finish, loose tolerances, slow cycle time

**Examples:** Engine blocks, machine tool bases, cylinder heads, pump housings, large gears

<details>
<summary>Discussion: Why is sand casting still widely used despite its limitations?</summary>

Sand casting remains popular because:

1. **Low startup cost** - A wood pattern costs hundreds vs. steel dies costing tens of thousands
2. **Flexibility** - Easy to modify patterns for design changes
3. **Size capability** - Can cast parts weighing several tons
4. **Material versatility** - Works with nearly any metal
5. **Short lead time** - Parts can be produced in days vs. weeks for die casting

For prototypes and low-volume production (< 500 parts), sand casting is often the most economical choice even with its lower quality.

</details>

### Die Casting

Die casting is used for producing metal parts where high precision is required. Molten metal is forced under high pressure into reusable metal dies, enabling high-volume production.

```{r video-die-casting, echo=FALSE, results='asis'}
embed_youtube("0oibUY8KUQM", "Die Casting Process")
```

**Process:**

1. Steel molds (dies) are machined to the part shape
2. Dies are mounted in a die casting machine
3. Molten metal is injected under high pressure (10-175 MPa)
4. Metal solidifies rapidly due to die cooling
5. Part is ejected and the cycle repeats

**Advantages:** Excellent dimensional accuracy, smooth surface finish, high production rates (up to 400 shots/hour)

**Disadvantages:** High die cost ($50,000-$500,000), limited to non-ferrous metals, porosity issues

**Examples:** Engine components, transmission housings, electronic enclosures, power tool housings

### Investment Casting (Lost-Wax)

Investment casting produces complex, precision parts by creating a wax pattern that is "invested" (surrounded) with ceramic slurry, then melted out to leave a mold cavity.

```{r video-investment-casting, echo=FALSE, results='asis'}
embed_youtube("QtxVdC7pBQM", "Investment Casting Process")
```

**Advantages:** Excellent surface finish, tight tolerances (±0.1%), complex geometries, any castable metal

**Disadvantages:** Expensive, slow cycle time, limited part size

**Examples:** Turbine blades, aerospace components, medical implants, jewelry, firearm components

<details>
<summary>Why do aerospace companies pay premium prices for investment casting?</summary>

Investment casting is essential for aerospace because:

1. **Complex internal passages** - Turbine blades have intricate cooling channels
2. **Superior metallurgy** - Controlled solidification produces better grain structure
3. **Near-net shape** - Minimizes machining of expensive superalloys
4. **Weight savings** - Can create hollow structures impossible with other methods
5. **Material options** - Works with titanium, Inconel, and other aerospace alloys

A single turbine blade might cost $1,000-$5,000, but there's no other process that can produce the required geometry and performance.

</details>

### Centrifugal Casting

Centrifugal casting spins molten metal in a rotating die, using centrifugal force to distribute the metal and float impurities to the inner surface.

**Applications:** Pipes, tubes, cylinder liners, bushings, rings

### Permanent Mold Casting

Uses reusable metal molds with gravity or low-pressure filling. Good for medium-volume production of aluminum and copper alloy parts.

**Applications:** Pistons, gear housings, wheels, pipe fittings

### Shell Molding

Uses resin-coated sand that forms a thin, strong shell around a heated pattern. Better dimensional accuracy than sand casting.

**Applications:** Gear housings, cylinder heads, connecting rods

------------------------------------------------------------------------

## 2. Molding Processes

**Molding** uses a rigid frame to shape hot liquid or ductile raw material. It is predominantly used for plastics and rubber products.

```{r molding-volume-chart, echo=FALSE, fig.align="center", fig.cap="Molding Process Selection by Production Volume", fig.width=9, fig.height=5}
molding_data <- data.frame(
  Process = c("Injection Molding", "Blow Molding", "Rotational Molding", "Thermoforming", "Extrusion"),
  Min_Volume = c(10000, 5000, 100, 500, 1000),
  Max_Volume = c(10000000, 1000000, 50000, 500000, 10000000),
  Typical = c(100000, 50000, 5000, 10000, 500000)
)

ggplot(molding_data, aes(x = reorder(Process, -Typical), y = Typical)) +
  geom_bar(stat = "identity", fill = "#3498DB", width = 0.6) +
  geom_errorbar(aes(ymin = Min_Volume, ymax = Max_Volume), width = 0.2, color = "#2C3E50") +
  scale_y_log10(labels = scales::comma) +
  labs(title = "Economical Production Volumes for Molding Processes",
       subtitle = "Error bars show typical volume range",
       x = "", y = "Annual Production Volume (log scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 20, hjust = 1)
  ) +
  coord_flip()
```

### Injection Molding

Injection molding is the most common process for manufacturing plastic parts. Plastic pellets are melted and injected under high pressure into a mold cavity.

```{r video-injection-molding, echo=FALSE, results='asis'}
embed_youtube("RMjtmsr3CqA", "Injection Molding Process")
```

**Process Cycle:**

1. **Clamping** - Mold halves are clamped together
2. **Injection** - Molten plastic is injected into the cavity
3. **Cooling** - Part solidifies in the mold
4. **Ejection** - Part is pushed out by ejector pins

**Cycle times:** 10-60 seconds typical, as fast as 2 seconds for small parts

**Examples:** Automotive dashboards, phone cases, medical devices, food containers, toys

```{r injection-molding-cost, echo=TRUE}
# Injection Molding Cost Analysis Example
# Comparing costs at different production volumes

mold_cost <- 50000  # Typical mold cost in dollars
material_cost_per_kg <- 2.50  # ABS plastic
part_weight_kg <- 0.1
cycle_time_sec <- 30
machine_rate_per_hour <- 75  # Machine + operator cost

# Calculate cost per part at different volumes
volumes <- c(1000, 10000, 50000, 100000, 500000)

cost_analysis <- data.frame(
  Volume = volumes,
  Mold_Per_Part = mold_cost / volumes,
  Material = part_weight_kg * material_cost_per_kg,
  Processing = (cycle_time_sec / 3600) * machine_rate_per_hour
)

cost_analysis$Total_Per_Part <- cost_analysis$Mold_Per_Part +
                                 cost_analysis$Material +
                                 cost_analysis$Processing

cat("Injection Molding Cost Breakdown:\n")
cat("================================\n")
for(i in 1:nrow(cost_analysis)) {
  cat(sprintf("At %s parts: $%.2f per part (Tooling: $%.2f, Material: $%.2f, Processing: $%.2f)\n",
              format(cost_analysis$Volume[i], big.mark=","),
              cost_analysis$Total_Per_Part[i],
              cost_analysis$Mold_Per_Part[i],
              cost_analysis$Material[i],
              cost_analysis$Processing[i]))
}
```

<details>
<summary>Discussion: When does injection molding become cost-effective vs. 3D printing?</summary>

**Break-even analysis:**

- 3D printing: ~$5-20 per part (no tooling, slow production)
- Injection molding: $50,000 tooling + $0.50-2 per part

**Break-even point:** Approximately 5,000-10,000 parts

**Rule of thumb:**
- < 500 parts: 3D printing or machining
- 500-5,000 parts: Consider soft tooling (aluminum molds)
- > 5,000 parts: Steel injection molds become economical

</details>

### Blow Molding

Blow molding creates hollow plastic parts by inflating a heated plastic tube (parison) inside a mold cavity.

**Types:**
- **Extrusion blow molding** - Parison extruded, then blown
- **Injection blow molding** - Preform injection molded, then blown
- **Stretch blow molding** - Preform stretched and blown (PET bottles)

**Examples:** Bottles, containers, fuel tanks, automotive ducts, toys

### Rotational Molding

Creates hollow parts by rotating a mold on two axes while heating, causing plastic powder to melt and coat the interior surface.

**Advantages:** Low tooling cost, no internal stresses, large hollow parts

**Examples:** Kayaks, playground equipment, tanks, coolers

### Thermoforming

Heats a plastic sheet until pliable, then forms it over a mold using vacuum, pressure, or mechanical force.

**Examples:** Packaging (blister packs), disposable cups, refrigerator liners

### Powder Metallurgy

Though involving metal, powder metallurgy uses molding principles. Metal powders are compacted in dies and sintered (heated below melting point) to fuse particles.

**Advantages:** Near-net shape, complex geometries, controlled porosity

**Examples:** Self-lubricating bearings, gears, filters, magnets

------------------------------------------------------------------------

## 3. Forming Processes

**Forming** uses mechanical forces (compression, tension, shear) to shape material without removing any material. This preserves material grain structure and can improve mechanical properties.

```{r forming-forces, echo=FALSE, fig.align="center", fig.cap="Types of Forming Forces", fig.width=9, fig.height=4}
# Visualization of forming force types
force_data <- data.frame(
  Type = c("Compression", "Tension", "Shear", "Bending"),
  Process = c("Forging, Rolling", "Drawing, Stretching", "Shearing, Blanking", "Brake Forming"),
  x = c(1, 2, 3, 4)
)

ggplot(force_data, aes(x = x, y = 1)) +
  geom_tile(aes(fill = Type), width = 0.8, height = 0.5, color = "black") +
  geom_text(aes(label = Type), y = 1.1, fontface = "bold", size = 4) +
  geom_text(aes(label = Process), y = 0.9, size = 3, color = "gray30") +
  scale_fill_manual(values = c("#E74C3C", "#F39C12", "#3498DB", "#2ECC71")) +
  labs(title = "Forming Process Classification by Primary Force Type") +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    legend.position = "none"
  )
```

### Forging

Forging uses compressive forces from hammers or presses to shape heated metal. Forged parts have superior strength due to aligned grain structure.

```{r video-forging, echo=FALSE, results='asis'}
embed_youtube("o4vQJnPwgiA", "Forging Process")
```

**Types of Forging:**
- **Open-die forging** - Simple shapes, large parts
- **Closed-die (impression) forging** - Complex shapes, high volume
- **Cold forging** - Room temperature, excellent finish
- **Hot forging** - Heated material, easier deformation

**Examples:** Crankshafts, connecting rods, gears, aircraft structural components, hand tools

<details>
<summary>Why are aircraft landing gear forgings so expensive?</summary>

Aircraft landing gear forgings can cost $100,000+ because:

1. **Material** - High-strength steel or titanium alloys
2. **Size** - Large parts require massive presses (50,000+ tons)
3. **Certification** - Extensive testing and documentation
4. **Grain flow** - Must be optimized for fatigue resistance
5. **NDT requirements** - 100% ultrasonic and magnetic particle inspection
6. **Low volume** - Only a few hundred per aircraft type per year

The forged grain structure provides 20-40% better fatigue life than equivalent castings or machined parts.

</details>

### Stamping

Stamping uses dies and presses to form, cut, or shape sheet metal. Modern stamping lines can produce 1,000+ parts per hour.

**Operations include:**
- **Blanking** - Cutting flat shapes from sheet
- **Piercing** - Punching holes
- **Drawing** - Forming cup shapes
- **Bending** - Creating angles
- **Coining** - Fine detail forming

**Examples:** Automotive body panels, appliance housings, electronic enclosures

### Bending

Bending uses press brakes to create angular shapes in sheet metal.

**Key formula:**
$$\text{Bend Allowance} = \frac{\pi \times A \times (R + K \times T)}{180}$$

Where: A = bend angle, R = inside radius, T = material thickness, K = K-factor (typically 0.3-0.5)

### Shearing

Shearing cuts sheet metal without chip formation using opposed cutting edges.

**Types:** Blanking, piercing, trimming, notching, slitting

### Rolling

Reduces material thickness by passing through rotating rolls. Used for sheet, plate, and structural shapes.

### Extrusion (Metal)

Forces material through a die to create constant cross-section profiles.

**Examples:** Aluminum window frames, heat sinks, structural shapes

------------------------------------------------------------------------

## 4. Machining Processes

**Machining** removes material using cutting tools to achieve precise dimensions and surface finishes. It is the most versatile manufacturing process but typically has higher per-part costs.

```{r machining-comparison, echo=FALSE, fig.align="center", fig.cap="Machining Process Capabilities", fig.width=10, fig.height=5}
machining_data <- data.frame(
  Process = c("Turning", "Milling", "Drilling", "Grinding", "EDM", "Laser Cutting"),
  Tolerance_um = c(25, 25, 50, 2, 5, 100),
  Surface_Ra = c(1.6, 1.6, 3.2, 0.2, 0.8, 3.2),
  Material_Removal = c(5, 4, 3, 1, 1, 2)  # Relative rate
)

machining_long <- machining_data %>%
  mutate(
    `Tolerance (μm)` = Tolerance_um,
    `Surface Finish (Ra μm)` = Surface_Ra
  ) %>%
  select(Process, `Tolerance (μm)`, `Surface Finish (Ra μm)`) %>%
  pivot_longer(cols = -Process, names_to = "Metric", values_to = "Value")

ggplot(machining_long, aes(x = Process, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  scale_fill_manual(values = c("#3498DB", "#E74C3C")) +
  facet_wrap(~Metric, scales = "free_y") +
  labs(title = "Machining Process Precision Comparison",
       subtitle = "Lower values indicate better precision",
       x = "", y = "Value") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

### Turning

Rotates the workpiece while a cutting tool moves to create cylindrical shapes.

**Performed on lathes or turning centers**

**Operations:** Facing, turning, boring, threading, grooving, parting

### Milling

Rotates a cutting tool while moving the workpiece to remove material. The most versatile machining process.

**Types:**
- **Peripheral milling** - Cutter axis parallel to surface
- **Face milling** - Cutter axis perpendicular to surface
- **End milling** - Creates slots, pockets, contours

### Drilling

Creates cylindrical holes using rotating drill bits.

**Related operations:** Reaming (sizing), boring (enlarging), tapping (threading)

### Grinding

Uses abrasive wheels for precision finishing. Achieves the finest tolerances and surface finishes.

**Types:** Surface grinding, cylindrical grinding, centerless grinding

### Electrical Discharge Machining (EDM)

Uses electrical sparks to erode conductive materials. Essential for hardened materials and complex geometries.

**Applications:** Injection mold cavities, dies, aerospace components

<details>
<summary>When should you use EDM instead of conventional machining?</summary>

**Choose EDM when:**

1. **Hardened materials** - Can machine 60+ HRC steels that would destroy cutting tools
2. **Complex internal features** - Wire EDM cuts intricate shapes
3. **Tight corners** - Can achieve sharp internal corners impossible with rotary tools
4. **Fragile parts** - No cutting forces (non-contact process)
5. **Thin walls** - No deflection from tool pressure

**Limitations:**
- Slow material removal (grams/hour vs. kg/hour for milling)
- Only works on conductive materials
- Recast layer may need removal
- Higher cost per part

</details>

------------------------------------------------------------------------

## 5. Joining Processes

**Joining** combines two or more components into a single assembly. Methods range from permanent (welding) to removable (fasteners).

```{r joining-comparison, echo=FALSE}
joining_data <- tibble(
  Process = c("Arc Welding", "Resistance Welding", "Laser Welding", "Brazing", "Soldering", "Adhesive Bonding", "Mechanical Fastening"),
  `Joint Strength` = c("Excellent", "Excellent", "Excellent", "Good", "Fair", "Good", "Excellent"),
  `Dissimilar Metals` = c("Limited", "Limited", "Limited", "Yes", "Yes", "Yes", "Yes"),
  `Automation` = c("Medium", "High", "High", "Medium", "High", "High", "High"),
  `Typical Applications` = c("Structural steel, pressure vessels", "Auto body panels, wire mesh", "Electronics, medical devices", "Carbide tooling, HVAC", "Electronics, plumbing", "Composites, glass, plastics", "All industries")
)

joining_data %>%
  kable(format = "html", caption = "Joining Process Comparison") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    full_width = TRUE
  ) %>%
  row_spec(0, bold = TRUE, background = "#2C3E50", color = "white") %>%
  column_spec(1, bold = TRUE, background = "#ECF0F1")
```

### Welding

Melts and fuses base materials together, often with added filler metal.

**Common Types:**
- **MIG (GMAW)** - Gas Metal Arc Welding - fast, easy automation
- **TIG (GTAW)** - Gas Tungsten Arc Welding - precise, clean welds
- **Stick (SMAW)** - Shielded Metal Arc Welding - versatile, field use
- **Resistance** - Uses electrical resistance heating
- **Laser/Electron Beam** - High precision, deep penetration

### Brazing and Soldering

Joins metals using a filler metal with a lower melting point than the base materials. The base materials do not melt.

- **Brazing** - Filler melts above 450°C (840°F)
- **Soldering** - Filler melts below 450°C

### Adhesive Bonding

Uses chemical adhesives to join materials. Essential for composites, glass, and dissimilar materials.

**Advantages:** Distributes stress, joins dissimilar materials, seals against fluids

**Disadvantages:** Surface preparation critical, limited temperature resistance, aging concerns

### Mechanical Fastening

Uses bolts, screws, rivets, or clips. Allows disassembly for maintenance.

------------------------------------------------------------------------

## 6. Surface Treatment and Coating

**Coating and surface treatment** processes protect parts from corrosion, wear, or environmental damage, and can improve appearance or electrical properties.

```{r coating-processes, echo=FALSE}
coating_data <- tibble(
  Process = c("Electroplating", "Anodizing", "Powder Coating", "Painting", "Hot Dip Galvanizing", "PVD/CVD", "Thermal Spray"),
  `Typical Thickness` = c("1-50 μm", "5-100 μm", "50-150 μm", "25-75 μm", "45-85 μm", "1-10 μm", "100-500 μm"),
  `Base Material` = c("Conductive metals", "Aluminum", "Any metal", "Any", "Steel", "Any", "Any"),
  `Primary Benefit` = c("Corrosion, appearance", "Corrosion, hardness", "Corrosion, appearance", "Appearance", "Corrosion", "Wear, hardness", "Wear, thermal barrier"),
  Industry = c("Automotive, electronics", "Aerospace, construction", "Appliances, furniture", "All", "Construction, automotive", "Tooling, aerospace", "Aerospace, energy")
)

coating_data %>%
  kable(format = "html", caption = "Surface Treatment and Coating Processes") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    full_width = TRUE
  ) %>%
  row_spec(0, bold = TRUE, background = "#2C3E50", color = "white") %>%
  column_spec(1, bold = TRUE, background = "#ECF0F1")
```

### Electroplating

Deposits metal coating using electrical current. Common coatings include chrome, nickel, zinc, gold, and silver.

### Anodizing

Electrochemically grows an oxide layer on aluminum. Provides corrosion resistance and can be dyed various colors.

### Powder Coating

Applies dry powder electrostatically, then cures with heat. Produces thick, durable, attractive finishes.

### Galvanizing

Coats steel with zinc for corrosion protection. Hot-dip galvanizing provides excellent outdoor durability.

------------------------------------------------------------------------

## 7. Additive Manufacturing (3D Printing)

**Additive manufacturing** builds parts layer-by-layer from digital models. It's the opposite of machining (subtractive manufacturing).

```{r am-processes, echo=FALSE, fig.align="center", fig.cap="Additive Manufacturing Process Comparison", fig.width=10, fig.height=6}
am_data <- data.frame(
  Process = c("FDM", "SLA", "SLS", "DMLS/SLM", "Binder Jetting", "PolyJet"),
  Material = c("Thermoplastics", "Photopolymers", "Nylon, Polymers", "Metals", "Metals, Sand", "Photopolymers"),
  Accuracy = c(2, 5, 3, 4, 3, 5),
  Speed = c(3, 2, 3, 1, 4, 2),
  Cost = c(5, 3, 2, 1, 3, 2),
  Strength = c(3, 2, 4, 5, 4, 2)
)

am_long <- am_data %>%
  select(-Material) %>%
  pivot_longer(cols = -Process, names_to = "Metric", values_to = "Score")

ggplot(am_long, aes(x = Metric, y = Score, fill = Process)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(breaks = 1:5, labels = c("Poor", "Fair", "Good", "Very Good", "Excellent")) +
  labs(title = "Additive Manufacturing Process Comparison",
       subtitle = "Comparing key characteristics across 3D printing technologies",
       x = "", y = "Rating") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```

```{r video-3d-printing, echo=FALSE, results='asis'}
  embed_youtube("Vx0Z6LplaMU", "How 3D Printing Works")
```


### FDM (Fused Deposition Modeling)

Extrudes thermoplastic filament through a heated nozzle. Most common desktop 3D printing technology.

**Materials:** PLA, ABS, PETG, Nylon, TPU

### SLA (Stereolithography)

Uses UV laser to cure liquid photopolymer resin. Excellent surface finish and detail.

### SLS (Selective Laser Sintering)

Fuses powder material with a laser. Produces strong functional parts without support structures.

### Metal 3D Printing (DMLS/SLM)

Fuses metal powder with high-power lasers. Used for aerospace, medical, and tooling applications.

<details>
<summary>When does metal 3D printing make sense vs. traditional manufacturing?</summary>

**Metal 3D printing is advantageous when:**

1. **Complex internal features** - Conformal cooling channels in molds
2. **Topology optimization** - Organic shapes that minimize weight
3. **Consolidation** - Combining multiple parts into one
4. **Low volume** - Too few parts to justify tooling
5. **Custom/one-off** - Medical implants, prototypes

**Traditional manufacturing is better when:**
- High volumes (> 100-1000 parts depending on complexity)
- Simple geometries
- Large parts (3D print beds are limited)
- Maximum strength required (forging still superior)

**Cost comparison:**
- Simple bracket: $5 machined vs. $50+ 3D printed
- Complex aerospace part: $10,000 machined vs. $3,000 3D printed

</details>

------------------------------------------------------------------------

## Industry Applications

```{r industry-table, echo=FALSE}
industry_examples <- tibble(
  `Process` = c("Die Casting", "Injection Molding", "Forging", "Stamping", "Machining", "Welding", "Powder Coating", "3D Printing"),
  `Automotive` = c("Engine blocks, transmission cases", "Dashboard, interior trim, connectors", "Crankshafts, connecting rods, gears", "Body panels, structural components", "Engine components, precision parts", "Body structure, exhaust systems", "Wheels, chassis components", "Prototypes, custom fixtures"),
  `Food Industry` = c("Equipment housings", "Packaging, containers, caps", "Mixer blades, pump impellers", "Cans, trays, lids", "Valves, fittings", "Tanks, piping, conveyors", "Equipment frames", "Custom tooling, fixtures"),
  `Defense/Aerospace` = c("Aircraft structural components", "Connectors, housings, switches", "Landing gear, turbine disks", "Aircraft skins, missile bodies", "Precision weapons components", "Aircraft structures, armor", "Military vehicles, equipment", "Turbine blades, brackets, tooling")
)

industry_examples %>%
  kable(format = "html", caption = "Manufacturing Process Applications by Industry") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    full_width = TRUE
  ) %>%
  row_spec(0, bold = TRUE, background = "#1A5276", color = "white") %>%
  column_spec(1, bold = TRUE, background = "#D4E6F1") %>%
  column_spec(2, background = "#FCF3CF") %>%
  column_spec(3, background = "#D5F5E3") %>%
  column_spec(4, background = "#FADBD8")
```

------------------------------------------------------------------------

## Sustainability in Manufacturing

Modern manufacturing must consider environmental impact. Here's how different processes compare:

```{r sustainability, echo=FALSE, fig.align="center", fig.cap="Environmental Considerations by Process Type", fig.width=10, fig.height=5}
sustainability_data <- data.frame(
  Process = c("Casting", "Molding", "Forming", "Machining", "Joining", "Additive"),
  Material_Waste = c(2, 1, 1, 4, 1, 2),
  Energy_Use = c(5, 3, 4, 2, 4, 3),
  Recyclability = c(5, 3, 5, 5, 4, 2)
)

sustainability_long <- sustainability_data %>%
  pivot_longer(cols = -Process, names_to = "Factor", values_to = "Score") %>%
  mutate(Factor = gsub("_", " ", Factor))

ggplot(sustainability_long, aes(x = Process, y = Score, fill = Factor)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  scale_fill_manual(values = c("#27AE60", "#E74C3C", "#3498DB")) +
  scale_y_continuous(breaks = 1:5, labels = c("Very Low", "Low", "Medium", "High", "Very High")) +
  labs(title = "Sustainability Comparison of Manufacturing Processes",
       subtitle = "Lower Material Waste & Energy Use is better; Higher Recyclability is better",
       x = "", y = "Rating") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```

<details>
<summary>Discussion: How can manufacturers reduce environmental impact?</summary>

**Strategies for sustainable manufacturing:**

1. **Near-net-shape processes** - Casting, forging, and molding reduce machining waste
2. **Closed-loop recycling** - Reprocess scrap material on-site
3. **Energy-efficient equipment** - Modern machines use 30-50% less energy
4. **Additive manufacturing** - Only uses material where needed
5. **Water-based coatings** - Replace solvent-based paints
6. **Lean manufacturing** - Reduces overproduction and waste
7. **Life cycle assessment** - Consider total environmental impact

**Example:** Aluminum die casting recycles 95%+ of sprues and runners, while machining from billet may waste 80% of material.

</details>

------------------------------------------------------------------------

## Review Questions

<details>
<summary>Question 1: Process Identification</summary>

**A company needs to produce 50,000 aluminum housings per year. Each housing is a complex shape with internal fins for heat dissipation. The surface must be smooth and dimensions must be within ±0.3mm. Which process would you recommend?**

**Answer:** **Die Casting** is the best choice because:
- High volume (50,000/year) justifies tooling cost
- Aluminum is ideal for die casting
- Complex internal features are possible
- ±0.3mm tolerance is achievable
- Good surface finish without secondary operations

Alternative: If volumes were lower (< 5,000), investment casting might be considered.

</details>

<details>
<summary>Question 2: Cost Comparison</summary>

**Calculate the break-even point between 3D printing and injection molding for a small plastic part:**
- 3D printing: $8 per part (no tooling)
- Injection molding: $25,000 mold + $0.50 per part

**Answer:**
Let n = number of parts

Break-even: $8n = $25,000 + $0.50n

$7.50n = $25,000

n = 3,333 parts

**Below 3,333 parts:** 3D printing is cheaper
**Above 3,333 parts:** Injection molding is cheaper

At 10,000 parts:
- 3D printing: $80,000
- Injection molding: $30,000

</details>

<details>
<summary>Question 3: Material Selection</summary>

**Match each material/product to the most appropriate manufacturing process:**

a) Stainless steel surgical instrument
b) Large plastic kayak
c) Titanium jet engine bracket
d) Aluminum beverage can

**Answers:**

a) **Investment casting** or **machining** - Complex shapes, high precision, excellent surface finish
b) **Rotational molding** - Large hollow plastic part, low tooling cost
c) **Additive manufacturing (DMLS)** or **forging** - Complex geometry, lightweight, strength critical
d) **Deep drawing (stamping)** - Very high volume, thin-wall cylinder

</details>

<details>
<summary>Question 4: Process Selection Scenario</summary>

**You need to manufacture a new product with these requirements:**
- Quantity: 200 units for initial market test
- Material: ABS plastic
- Size: 150mm × 100mm × 50mm
- Tolerance: ±0.5mm
- Surface: Smooth, will be painted

**What process would you recommend and why?**

**Answer:** **3D Printing (FDM or SLA)** would be most appropriate because:

1. **Low quantity (200)** - Injection molding tooling ($30,000+) not justified
2. **Market test phase** - Design may change based on feedback
3. **Quick turnaround** - Parts in days vs. weeks for tooling
4. **Tolerances achievable** - ±0.5mm is within FDM capabilities
5. **Post-processing acceptable** - Parts will be painted anyway

**Cost comparison:**
- FDM: 200 × $15 = $3,000
- Injection molding: $30,000 + (200 × $1) = $30,200

If the product succeeds and volume increases to 10,000+, transition to injection molding.

</details>

<details>
<summary>Question 5: Hybrid Manufacturing</summary>

**What is hybrid manufacturing and when is it used?**

**Answer:** **Hybrid manufacturing** combines additive and subtractive processes in a single machine or workflow.

**Common combinations:**
- 3D print near-net shape → CNC machine critical surfaces
- Cast blank → Machine precision features
- Forge rough shape → Machine final dimensions

**Benefits:**
- Reduced material waste vs. pure machining
- Better precision than pure additive
- Faster than pure machining for complex parts

**Example:** Aerospace companies 3D print titanium brackets, then machine mounting surfaces to achieve tight tolerances. This can reduce material waste by 90% compared to machining from billet.

</details>

------------------------------------------------------------------------

## Summary

```{r summary-table-mfg, echo=FALSE}
summary_data <- tibble(
  `Process Category` = c("Casting", "Molding", "Forming", "Machining", "Joining", "Coating", "Additive"),
  `Key Principle` = c("Pour liquid into mold", "Shape heated/softened material", "Deform solid with force", "Remove material with tools", "Combine multiple parts", "Apply protective layer", "Build layer by layer"),
  `Best For` = c("Complex metal shapes, medium-high volume", "Plastic parts, very high volume", "Strong metal parts, high volume", "Precision parts, any material", "Assembly of components", "Corrosion/wear protection", "Prototypes, complex geometry"),
  `Typical Industries` = c("Automotive, aerospace", "Consumer products, packaging", "Automotive, construction", "All industries", "All industries", "All industries", "Aerospace, medical, prototyping")
)

summary_data %>%
  kable(format = "html", caption = "Manufacturing Processes Summary") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "bordered"),
    full_width = TRUE
  ) %>%
  row_spec(0, bold = TRUE, background = "#1A5276", color = "white") %>%
  column_spec(1, bold = TRUE)
```

### Key Takeaways

1. **No single process is best for everything** - Selection depends on material, volume, complexity, and cost constraints
2. **Volume drives economics** - High tooling costs require high volumes to amortize
3. **Consider the complete supply chain** - Secondary operations, assembly, and finishing affect total cost
4. **Sustainability matters** - Modern manufacturing must minimize waste and energy use
5. **Hybrid approaches are growing** - Combining processes optimizes cost and performance

------------------------------------------------------------------------

## References

1. Groover, M.P. (2020). *Fundamentals of Modern Manufacturing* (7th ed.). Wiley.
2. Kalpakjian, S., & Schmid, S.R. (2020). *Manufacturing Engineering and Technology* (8th ed.). Pearson.
3. DeGarmo, E.P., Black, J.T., & Kohser, R.A. (2019). *DeGarmo's Materials and Processes in Manufacturing* (13th ed.). Wiley.
4. Afzal, M. "Types of Manufacturing Processes." *Engineering Articles*.
