# Quality Management Systems
```{r setup-14, include=FALSE}
# include the helpers.R file for pdf output of youtube on pdfs as a clickable link
source("R/helpers.R")
# include the required packages to make sure the page can be built.
source("R/required_packages.R")
```
---------------------

## Learning Objectives

After completing this chapter, you will be able to:

1. Define quality management systems and explain their purpose in manufacturing
2. Describe the structure and requirements of ISO 9001:2015
3. Identify industry-specific quality standards (IATF 16949, AS9100, ISO 22000)
4. Explain the documentation hierarchy and control requirements
5. Conduct internal audits using systematic methodologies
6. Implement effective corrective and preventive action (CAPA) processes
7. Understand management review and continuous improvement requirements
8. Integrate quality tools (SPC, PFMEA, MSA) within a QMS framework

---

## Introduction to Quality Management Systems

A **Quality Management System (QMS)** is a formalized system that documents processes, procedures, and responsibilities for achieving quality policies and objectives. It coordinates and directs an organization's activities to meet customer and regulatory requirements while continuously improving effectiveness.

### Why Implement a QMS?

```{r qms-benefits, echo=FALSE, fig.width=12, fig.height=7}
benefits <- data.frame(
  Category = c("Customer", "Operational", "Financial", "Strategic", "Regulatory", "Cultural"),
  Benefit = c("Customer Satisfaction", "Process Consistency", "Cost Reduction",
              "Market Access", "Compliance", "Employee Engagement"),
  Description = c(
    "Meet customer requirements consistently; reduce complaints",
    "Standardized processes; reduced variation; repeatable results",
    "Less scrap, rework, warranty; improved efficiency",
    "Required for many contracts; competitive advantage",
    "Meet regulatory requirements; reduce liability",
    "Clear expectations; empowerment; pride in quality"
  ),
  Impact = c(90, 85, 80, 88, 75, 70)
)

ggplot(benefits, aes(x = reorder(Category, Impact), y = Impact, fill = Category)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_text(aes(label = Benefit), hjust = -0.1, size = 3.5, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = c("#e74c3c", "#3498db", "#27ae60",
                               "#f39c12", "#9b59b6", "#1abc9c"), guide = "none") +
  scale_y_continuous(limits = c(0, 120)) +
  labs(title = "Benefits of Implementing a Quality Management System",
       subtitle = "Key drivers for QMS adoption",
       x = "", y = "Relative Impact Score") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.y = element_text(size = 11, face = "bold")
  )
```

### The Evolution of Quality Management

```{r quality-evolution, echo=FALSE}
evolution <- data.frame(
  Era = c("Pre-1900s", "1900-1940s", "1940-1960s", "1960-1980s", "1980-2000s", "2000s-Present"),
  Approach = c("Craftsmanship", "Inspection", "Statistical Quality Control",
               "Quality Assurance", "Total Quality Management", "Integrated Management"),
  Focus = c(
    "Individual skill; master craftsmen",
    "Sort good from bad; end-of-line inspection",
    "Control charts; sampling; Shewhart, Deming",
    "Prevention; systems approach; documented procedures",
    "Company-wide quality; customer focus; continuous improvement",
    "Risk-based thinking; integration with business strategy; digital quality"
  ),
  Key_Figure = c("Guilds", "Taylor", "Shewhart, Deming", "Juran, Feigenbaum",
                 "Crosby, ISO 9000", "ISO 9001:2015, Industry 4.0")
)

kable(evolution, col.names = c("Era", "Approach", "Focus", "Key Figures/Standards"),
      caption = "Evolution of Quality Management") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "12%") %>%
  row_spec(6, background = "#d5f5e3")
```

### Quality Management Principles

ISO 9000 defines seven quality management principles that form the foundation of all QMS standards:

```{r qm-principles, echo=FALSE, fig.width=12, fig.height=8}
principles <- data.frame(
  principle = c("Customer Focus", "Leadership", "Engagement of People",
                "Process Approach", "Improvement", "Evidence-based\nDecision Making",
                "Relationship\nManagement"),
  number = 1:7,
  description = c(
    "Understand and meet customer needs",
    "Create unity of purpose and direction",
    "Competent, empowered people at all levels",
    "Manage activities as interconnected processes",
    "Ongoing focus on improvement",
    "Decisions based on data analysis",
    "Manage relationships with interested parties"
  ),
  x = c(4, 1, 7, 2.5, 5.5, 2.5, 5.5),
  y = c(4, 2.5, 2.5, 1, 1, -0.5, -0.5)
)

ggplot(principles, aes(x = x, y = y)) +
  geom_point(size = 22, color = "#3498db") +
  geom_text(aes(label = number), color = "white", size = 8, fontface = "bold") +
  geom_text(aes(label = principle, y = y - 0.7), size = 3, fontface = "bold", lineheight = 0.9) +
  # Customer focus at top
  annotate("segment", x = 4, xend = 1, y = 3.5, yend = 2.8,
           arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +
  annotate("segment", x = 4, xend = 7, y = 3.5, yend = 2.8,
           arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +
  annotate("segment", x = 4, xend = 2.5, y = 3.5, yend = 1.5,
           arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +
  annotate("segment", x = 4, xend = 5.5, y = 3.5, yend = 1.5,
           arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +
  coord_cartesian(xlim = c(0, 8), ylim = c(-1.5, 5)) +
  labs(title = "Seven Quality Management Principles (ISO 9000:2015)",
       subtitle = "Customer Focus is the primary principle; others support it") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

---

## ISO 9001:2015 Structure

**ISO 9001:2015** is the international standard for quality management systems. It uses the High-Level Structure (HLS) common to all ISO management system standards.

### The Ten Clauses

```{r iso-clauses, echo=FALSE, fig.width=12, fig.height=9}
clauses <- data.frame(
  clause = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"),
  title = c("Scope", "Normative References", "Terms and Definitions",
            "Context of the Organization", "Leadership", "Planning",
            "Support", "Operation", "Performance Evaluation", "Improvement"),
  description = c(
    "Defines applicability of the standard",
    "References to ISO 9000:2015",
    "Definitions used in the standard",
    "Understanding the organization and its context",
    "Top management commitment and policy",
    "Actions to address risks and opportunities",
    "Resources, competence, awareness, communication",
    "Operational planning and control",
    "Monitoring, measurement, analysis, evaluation",
    "Nonconformity, corrective action, continual improvement"
  ),
  type = c("Intro", "Intro", "Intro", "Core", "Core", "Core",
           "Core", "Core", "Core", "Core"),
  y = c(10, 9, 8, 7, 6, 5, 4, 3, 2, 1)
)

ggplot(clauses, aes(x = 1, y = y)) +
  geom_tile(aes(fill = type), width = 4, height = 0.8, color = "white", size = 1) +
  geom_text(aes(x = -0.3, label = clause), fontface = "bold", size = 5, color = "#2c3e50") +
  geom_text(aes(x = 0.3, label = title), fontface = "bold", size = 3.5, hjust = 0, color = "white") +
  geom_text(aes(x = 2.8, label = description), size = 2.8, hjust = 1, color = "white") +
  scale_fill_manual(values = c("Intro" = "#95a5a6", "Core" = "#3498db"), guide = "none") +
  coord_cartesian(xlim = c(-1, 3.5)) +
  labs(title = "ISO 9001:2015 Structure - The Ten Clauses",
       subtitle = "Clauses 4-10 contain the auditable requirements") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### The Process Approach and PDCA

ISO 9001:2015 emphasizes the **process approach** combined with the **Plan-Do-Check-Act (PDCA)** cycle:

```{r pdca-cycle, echo=FALSE, fig.width=11, fig.height=8}
ggplot() +
  # PDCA circle

  annotate("path",
           x = 5 + 3 * cos(seq(0, 2*pi, length.out = 100)),
           y = 5 + 3 * sin(seq(0, 2*pi, length.out = 100)),
           color = "gray70", size = 2) +

  # Plan quadrant
  annotate("path",
           x = 5 + 3 * cos(seq(pi/2, pi, length.out = 50)),
           y = 5 + 3 * sin(seq(pi/2, pi, length.out = 50)),
           color = "#3498db", size = 4) +
  annotate("point", x = 5 - 2.1, y = 5 + 2.1, size = 20, color = "#3498db") +
  annotate("text", x = 5 - 2.1, y = 5 + 2.1, label = "PLAN", color = "white",
           fontface = "bold", size = 5) +
  annotate("text", x = 2.2, y = 7.5, label = "Clause 4: Context\nClause 5: Leadership\nClause 6: Planning",
           size = 3, hjust = 0, lineheight = 0.9) +

  # Do quadrant
  annotate("path",
           x = 5 + 3 * cos(seq(0, pi/2, length.out = 50)),
           y = 5 + 3 * sin(seq(0, pi/2, length.out = 50)),
           color = "#27ae60", size = 4) +
  annotate("point", x = 5 + 2.1, y = 5 + 2.1, size = 20, color = "#27ae60") +
  annotate("text", x = 5 + 2.1, y = 5 + 2.1, label = "DO", color = "white",
           fontface = "bold", size = 5) +
  annotate("text", x = 7.8, y = 7.5, label = "Clause 7: Support\nClause 8: Operation",
           size = 3, hjust = 1, lineheight = 0.9) +

  # Check quadrant
  annotate("path",
           x = 5 + 3 * cos(seq(-pi/2, 0, length.out = 50)),
           y = 5 + 3 * sin(seq(-pi/2, 0, length.out = 50)),
           color = "#f39c12", size = 4) +
  annotate("point", x = 5 + 2.1, y = 5 - 2.1, size = 20, color = "#f39c12") +
  annotate("text", x = 5 + 2.1, y = 5 - 2.1, label = "CHECK", color = "white",
           fontface = "bold", size = 4.5) +
  annotate("text", x = 7.8, y = 2.5, label = "Clause 9: Performance\nEvaluation",
           size = 3, hjust = 1, lineheight = 0.9) +

  # Act quadrant
  annotate("path",
           x = 5 + 3 * cos(seq(pi, 3*pi/2, length.out = 50)),
           y = 5 + 3 * sin(seq(pi, 3*pi/2, length.out = 50)),
           color = "#e74c3c", size = 4) +
  annotate("point", x = 5 - 2.1, y = 5 - 2.1, size = 20, color = "#e74c3c") +
  annotate("text", x = 5 - 2.1, y = 5 - 2.1, label = "ACT", color = "white",
           fontface = "bold", size = 5) +
  annotate("text", x = 2.2, y = 2.5, label = "Clause 10: Improvement",
           size = 3, hjust = 0, lineheight = 0.9) +

  # Center
  annotate("text", x = 5, y = 5, label = "QMS", fontface = "bold", size = 6) +

  # Arrows showing cycle direction
  annotate("segment", x = 3, xend = 2.5, y = 6.5, yend = 5.5,
           arrow = arrow(length = unit(0.3, "cm")), size = 1.2, color = "gray40") +
  annotate("segment", x = 7, xend = 7.5, y = 6.5, yend = 5.5,
           arrow = arrow(length = unit(0.3, "cm")), size = 1.2, color = "gray40") +
  annotate("segment", x = 7.5, xend = 7, y = 4.5, yend = 3.5,
           arrow = arrow(length = unit(0.3, "cm")), size = 1.2, color = "gray40") +
  annotate("segment", x = 2.5, xend = 3, y = 4.5, yend = 3.5,
           arrow = arrow(length = unit(0.3, "cm")), size = 1.2, color = "gray40") +

  coord_cartesian(xlim = c(1, 9), ylim = c(1, 9)) +
  labs(title = "PDCA Cycle and ISO 9001:2015 Clauses",
       subtitle = "The continuous improvement cycle embedded in the standard") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### Risk-Based Thinking

ISO 9001:2015 introduced **risk-based thinking** throughout the QMS:

```{r risk-based, echo=FALSE}
risk_table <- data.frame(
  Clause = c("4.1", "4.2", "6.1", "8.1", "9.1", "10.2"),
  Requirement = c(
    "Context of the Organization",
    "Interested Parties",
    "Actions to Address Risks and Opportunities",
    "Operational Planning and Control",
    "Monitoring and Measurement",
    "Nonconformity and Corrective Action"
  ),
  Risk_Application = c(
    "Identify internal/external issues that could affect QMS outcomes",
    "Understand needs/expectations that could impact quality",
    "Plan actions to address risks and opportunities; integrate into processes",
    "Control processes considering risks identified in planning",
    "Monitor effectiveness of actions taken to address risks",
    "Analyze nonconformities; update risks/opportunities as needed"
  )
)

kable(risk_table, caption = "Risk-Based Thinking in ISO 9001:2015") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "10%") %>%
  column_spec(2, width = "25%")
```

<details><summary>**Understanding Risk-Based Thinking**</summary>

**Risk-based thinking** doesn't require formal risk management (like FMEA for every process), but rather:

1. **Consider risks** when designing and implementing the QMS
2. **Preventive action** is built into the system, not a separate activity
3. **Proportional response** - more rigorous controls for higher-risk processes
4. **Opportunities** are also considered (not just negative risks)

**Example:** A food manufacturer identifies "allergen cross-contamination" as a significant risk. Risk-based thinking leads them to:
- Dedicated production lines for allergen-containing products
- Enhanced cleaning verification procedures
- Staff training on allergen awareness
- Increased inspection frequency after changeovers

The extent of controls is proportional to the risk level.
</details>

---

## Industry-Specific Quality Standards

Different industries have developed standards that build upon ISO 9001 with additional sector-specific requirements.

### Comparison of Major Standards

```{r standards-comparison, echo=FALSE, fig.width=12, fig.height=7}
standards <- data.frame(
  Standard = c("ISO 9001:2015", "IATF 16949:2016", "AS9100D", "ISO 22000:2018"),
  Industry = c("General", "Automotive", "Aerospace/Defense", "Food Safety"),
  Base = c("Standalone", "ISO 9001 + Automotive", "ISO 9001 + Aerospace", "ISO 9001 + HACCP"),
  x = c(1, 2, 3, 4),
  y = c(1, 1, 1, 1)
)

ggplot(standards, aes(x = x, y = y)) +
  geom_tile(aes(fill = Standard), width = 0.9, height = 1.5, color = "white", size = 2) +
  geom_text(aes(label = Standard), fontface = "bold", size = 4, color = "white", vjust = -1) +
  geom_text(aes(label = Industry), size = 3.5, color = "white") +
  geom_text(aes(label = Base), size = 2.8, color = "white", vjust = 2, lineheight = 0.9) +
  scale_fill_manual(values = c("#3498db", "#e74c3c", "#2c3e50", "#27ae60"), guide = "none") +
  labs(title = "Major Quality Management System Standards",
       subtitle = "Industry-specific standards build upon ISO 9001") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### IATF 16949:2016 (Automotive)

**IATF 16949** is the automotive quality standard, required by most major OEMs (Ford, GM, Toyota, VW, etc.).

```{r iatf-additions, echo=FALSE}
iatf_requirements <- data.frame(
  Area = c("Product Safety", "APQP/PPAP", "Control Plans", "MSA", "SPC",
           "Supplier Quality", "Warranty Management", "Manufacturing Feasibility"),
  ISO_9001 = c("Implied", "Not specified", "Not specified", "General requirement",
               "General requirement", "Basic purchasing", "Complaint handling", "Review of requirements"),
  IATF_16949 = c(
    "Explicit requirements for product safety processes",
    "APQP phases required; PPAP submission mandatory",
    "Control plans required for all parts; specific format",
    "MSA studies required per AIAG MSA manual",
    "SPC for all special characteristics",
    "Supplier development; second-party audits; IATF certification flow-down",
    "NTF analysis; warranty data analysis; field return analysis",
    "Documented manufacturing feasibility review"
  )
)

kable(iatf_requirements, col.names = c("Requirement Area", "ISO 9001", "IATF 16949 Addition"),
      caption = "IATF 16949 Additional Requirements Beyond ISO 9001") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "18%") %>%
  column_spec(2, background = "#fadbd8") %>%
  column_spec(3, background = "#d5f5e3")
```

### Automotive Core Tools

IATF 16949 requires the use of **five core tools**:

```{r core-tools, echo=FALSE, fig.width=12, fig.height=6}
core_tools <- data.frame(
  tool = c("APQP", "PPAP", "FMEA", "MSA", "SPC"),
  name = c("Advanced Product\nQuality Planning",
           "Production Part\nApproval Process",
           "Failure Mode &\nEffects Analysis",
           "Measurement\nSystems Analysis",
           "Statistical\nProcess Control"),
  purpose = c("Plan and define quality program",
              "Demonstrate production capability",
              "Identify and mitigate risks",
              "Ensure measurement validity",
              "Monitor and control processes"),
  x = 1:5
)

ggplot(core_tools, aes(x = x, y = 1)) +
  geom_tile(aes(fill = tool), width = 0.85, height = 1.5, color = "white", size = 1.5) +
  geom_text(aes(label = tool), fontface = "bold", size = 5, color = "white", vjust = -1) +
  geom_text(aes(label = name), size = 3, color = "white", lineheight = 0.9) +
  geom_text(aes(label = purpose, y = 0.3), size = 2.5, color = "white", lineheight = 0.9) +
  scale_fill_manual(values = c("#e74c3c", "#3498db", "#27ae60", "#f39c12", "#9b59b6"),
                    guide = "none") +
  labs(title = "AIAG Automotive Core Tools",
       subtitle = "Five essential tools required by IATF 16949") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### AS9100D (Aerospace/Defense)

**AS9100D** adds aerospace-specific requirements to ISO 9001:

```{r as9100-additions, echo=FALSE}
as9100_req <- data.frame(
  Area = c("Configuration Management", "First Article Inspection",
           "Counterfeit Parts Prevention", "Special Processes",
           "Product Safety", "Risk Management", "Human Factors"),
  Requirement = c(
    "Control product configuration throughout lifecycle; change management",
    "FAI per AS9102; documented verification of first production parts",
    "Controls to prevent counterfeit parts entering supply chain",
    "NADCAP accreditation for special processes (welding, heat treat, NDT)",
    "Product safety requirements; reporting of unsafe conditions",
    "Explicit risk management process (often using AS/NZS 4360 or similar)",
    "Consideration of human factors in process design"
  )
)

kable(as9100_req, caption = "AS9100D Key Additional Requirements") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "25%")
```

### ISO 22000:2018 (Food Safety)

**ISO 22000** integrates quality management with **HACCP** (Hazard Analysis Critical Control Points):

```{r iso22000-structure, echo=FALSE}
iso22000_elements <- data.frame(
  Element = c("Prerequisite Programs (PRPs)", "Operational PRPs (OPRPs)",
              "HACCP Plan", "Traceability", "Emergency Preparedness",
              "Validation and Verification"),
  Description = c(
    "Basic hygiene conditions: cleaning, pest control, personnel hygiene, facility design",
    "PRPs essential to control specific identified hazards",
    "Critical Control Points (CCPs) for significant hazards; critical limits; monitoring",
    "One-step-back, one-step-forward traceability; recall procedures",
    "Procedures for food safety emergencies; recall/withdrawal",
    "Validate control measures; verify HACCP plan effectiveness"
  ),
  Example = c(
    "Sanitation schedule, hand washing stations, pest control contract",
    "Metal detector for physical hazard control",
    "Pasteurization temperature ≥72°C for 15 seconds",
    "Lot coding, supplier records, distribution records",
    "Product recall procedure, mock recall exercises",
    "Challenge studies, environmental monitoring, internal audits"
  )
)

kable(iso22000_elements, caption = "ISO 22000 Food Safety Management Elements") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "22%")
```

---

## Documentation Requirements

A QMS requires documented information to ensure consistency, provide evidence, and enable improvement.

### Documentation Hierarchy

```{r doc-hierarchy, echo=FALSE, fig.width=11, fig.height=8}
ggplot() +
  # Level 1 - Policy
  annotate("polygon", x = c(5, 3, 7), y = c(7, 5.5, 5.5),
           fill = "#e74c3c", color = "black") +
  annotate("text", x = 5, y = 6.2, label = "Level 1\nQuality Policy\n& Objectives",
           color = "white", fontface = "bold", size = 3.5, lineheight = 0.9) +

  # Level 2 - Manual/Procedures
  annotate("polygon", x = c(5, 2, 8), y = c(5.3, 3.5, 3.5),
           fill = "#f39c12", color = "black") +
  annotate("text", x = 5, y = 4.3, label = "Level 2\nQuality Manual & Procedures\nWhat we do and who is responsible",
           color = "white", fontface = "bold", size = 3, lineheight = 0.9) +

  # Level 3 - Work Instructions
  annotate("polygon", x = c(5, 1, 9), y = c(3.3, 1.5, 1.5),
           fill = "#3498db", color = "black") +
  annotate("text", x = 5, y = 2.3, label = "Level 3\nWork Instructions, SOPs, Specifications\nHow we do it - step by step details",
           color = "white", fontface = "bold", size = 3, lineheight = 0.9) +

  # Level 4 - Records
  annotate("rect", xmin = 0.5, xmax = 9.5, ymin = 0.3, ymax = 1.3,
           fill = "#27ae60", color = "black") +
  annotate("text", x = 5, y = 0.8, label = "Level 4: Records & Forms - Evidence that we did it",
           color = "white", fontface = "bold", size = 3.5) +

  # Examples on the right
  annotate("text", x = 9, y = 6.2, label = "• Quality Policy statement\n• Quality objectives",
           hjust = 0, size = 2.8, lineheight = 0.9) +
  annotate("text", x = 9, y = 4.3, label = "• Document control procedure\n• Corrective action procedure\n• Internal audit procedure",
           hjust = 0, size = 2.8, lineheight = 0.9) +
  annotate("text", x = 9, y = 2.3, label = "• Setup instructions\n• Inspection checklists\n• Calibration procedures",
           hjust = 0, size = 2.8, lineheight = 0.9) +
  annotate("text", x = 9, y = 0.8, label = "• Inspection records • Training records • Audit reports",
           hjust = 0, size = 2.8) +

  coord_cartesian(xlim = c(0, 13), ylim = c(0, 7.5)) +
  labs(title = "Quality Documentation Hierarchy",
       subtitle = "From policy (strategic) to records (evidence)") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### Document Control Requirements

```{r doc-control, echo=FALSE}
doc_control <- data.frame(
  Requirement = c("Approval", "Review & Update", "Identification",
                  "Availability", "Protection", "Obsolete Control",
                  "External Documents", "Record Retention"),
  Description = c(
    "Documents reviewed and approved for adequacy before issue",
    "Reviewed, updated as necessary, and re-approved",
    "Identified with title, date, revision, author",
    "Relevant versions available at points of use",
    "Protected from loss, damage, unauthorized changes",
    "Obsolete documents identified and prevented from unintended use",
    "External documents identified and distribution controlled",
    "Records retained for specified periods; protected; retrievable"
  ),
  Implementation = c(
    "Approval signatures or electronic workflow",
    "Periodic review schedule; change request process",
    "Document numbering system; revision control",
    "Controlled copies at workstations; electronic access",
    "Backup systems; access controls; original storage",
    "Stamp 'OBSOLETE'; remove from use; archive",
    "List of external documents; master copy control",
    "Retention matrix; secure storage; indexing"
  )
)

kable(doc_control, caption = "Document Control Requirements (ISO 9001 Clause 7.5)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "18%")
```

### Documented Information Required by ISO 9001:2015

```{r mandatory-docs, echo=FALSE}
mandatory <- data.frame(
  Type = c(rep("Documents (shall be maintained)", 7),
           rep("Records (shall be retained)", 10)),
  Clause = c("4.3", "5.2", "6.2", "7.1.5", "7.2", "8.1", "8.5.1",
             "7.1.5.1", "7.2", "8.2.3.2", "8.3.3", "8.3.4", "8.3.5", "8.3.6",
             "8.5.2", "8.6", "9.1.1"),
  Requirement = c(
    "Scope of the QMS",
    "Quality policy",
    "Quality objectives",
    "Monitoring and measuring resources (calibration)",
    "Evidence of competence",
    "Evidence of conformity to processes",
    "Controlled conditions for production",
    "Calibration/verification results",
    "Training records",
    "Results of review of requirements",
    "Design and development inputs",
    "Design and development controls",
    "Design and development outputs",
    "Design and development changes",
    "Traceability requirements",
    "Release of products/services",
    "Results of monitoring and measurement"
  )
)

kable(mandatory, caption = "Documented Information Required by ISO 9001:2015") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "25%") %>%
  pack_rows("Documents (shall be maintained)", 1, 7, background = "#ebf5fb") %>%
  pack_rows("Records (shall be retained)", 8, 17, background = "#eafaf1")
```

---

## Internal Auditing

**Internal audits** are systematic, independent evaluations of the QMS to determine if it conforms to requirements and is effectively implemented.

### The Audit Process

```{r audit-process, echo=FALSE, fig.width=12, fig.height=6}
audit_steps <- data.frame(
  step = 1:7,
  name = c("Planning", "Preparation", "Opening Meeting",
           "Audit Execution", "Closing Meeting", "Reporting", "Follow-up"),
  description = c(
    "Schedule audits; assign auditors",
    "Review documents; prepare checklist",
    "Explain purpose, scope, methods",
    "Collect evidence; interview; observe",
    "Present findings; discuss",
    "Issue audit report",
    "Verify corrective actions"
  ),
  x = 1:7
)

ggplot(audit_steps, aes(x = x, y = 1)) +
  geom_point(size = 15, color = "#3498db") +
  geom_text(aes(label = step), color = "white", fontface = "bold", size = 5) +
  geom_text(aes(label = name, y = 0.5), size = 3, fontface = "bold", angle = 45, hjust = 1) +
  geom_text(aes(label = description, y = 1.5), size = 2.5, lineheight = 0.9) +
  geom_segment(aes(x = x + 0.3, xend = x + 0.7, y = 1, yend = 1),
               data = audit_steps[1:6,],
               arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +
  scale_x_continuous(limits = c(0.5, 7.5)) +
  scale_y_continuous(limits = c(0, 2)) +
  labs(title = "Internal Audit Process Flow",
       subtitle = "Systematic approach to evaluating QMS effectiveness") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### Audit Checklist Example

```{r audit-checklist, echo=FALSE}
checklist <- data.frame(
  Clause = c("7.1.5", "7.1.5", "7.1.5", "7.1.5", "7.1.5.2"),
  Requirement = c(
    "Monitoring and measuring resources suitable for the type of activities",
    "Resources maintained to ensure continued fitness for purpose",
    "Documented information on fitness for purpose retained",
    "Measuring equipment identified to determine status",
    "Measurement traceability when required"
  ),
  Question = c(
    "How do you determine what measuring equipment is needed?",
    "What maintenance is performed on measuring equipment?",
    "Can you show me calibration records for this micrometer?",
    "How can I tell if this gauge is currently calibrated?",
    "What standards are these gauges calibrated against?"
  ),
  Evidence = c("Gauge selection procedure", "PM records", "Calibration certificate",
               "Calibration stickers/tags", "Traceability certificates")
)

kable(checklist, caption = "Sample Audit Checklist: Clause 7.1.5 Monitoring and Measuring Resources") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "8%")
```

### Types of Audit Findings

```{r finding-types, echo=FALSE}
findings <- data.frame(
  Type = c("Major Nonconformity", "Minor Nonconformity", "Observation",
           "Opportunity for Improvement"),
  Definition = c(
    "Absence or total breakdown of a system; significant risk to product/customer",
    "Single lapse or isolated incident; system exists but not fully effective",
    "Weakness that could lead to nonconformity if not addressed",
    "Suggestion for enhancement; not a nonconformity"
  ),
  Example = c(
    "No calibration program exists; all gauges are uncontrolled",
    "3 of 50 gauges found past due for calibration",
    "Calibration procedure doesn't specify what to do if gauge fails",
    "Consider using electronic calibration records for faster retrieval"
  ),
  Action_Required = c(
    "Immediate corrective action; root cause analysis; may affect certification",
    "Corrective action required; typically 30-90 days",
    "No formal corrective action; monitor situation",
    "Consider for improvement; no requirement"
  )
)

kable(findings, col.names = c("Finding Type", "Definition", "Example", "Action Required"),
      caption = "Types of Audit Findings") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "18%") %>%
  row_spec(1, background = "#f8d7da") %>%
  row_spec(2, background = "#fff3cd") %>%
  row_spec(3, background = "#d1ecf1") %>%
  row_spec(4, background = "#d4edda")
```

### Auditor Competence

```{r auditor-competence, echo=FALSE}
auditor_skills <- data.frame(
  Competency = c("Knowledge of Standards", "Audit Techniques", "Industry Knowledge",
                 "Communication Skills", "Objectivity", "Professional Judgment"),
  Description = c(
    "Understanding of ISO 9001 and applicable industry standards",
    "Planning, interviewing, evidence collection, reporting",
    "Understanding of processes, products, and industry practices",
    "Active listening, clear questioning, professional writing",
    "Independence from area being audited; impartial",
    "Ability to evaluate significance of findings"
  ),
  How_Developed = c(
    "Training courses; self-study; certification (CQA, Lead Auditor)",
    "Auditor training; shadowing experienced auditors",
    "Work experience; process knowledge; technical training",
    "Practice; feedback; soft skills training",
    "Audit program design; rotation of assignments",
    "Experience; mentoring; calibration with other auditors"
  )
)

kable(auditor_skills, caption = "Internal Auditor Competency Requirements") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "20%")
```

---

## Corrective and Preventive Action (CAPA)

**Corrective action** eliminates the cause of a detected nonconformity to prevent recurrence. **Preventive action** eliminates the cause of a potential nonconformity to prevent occurrence.

### The CAPA Process

```{r capa-process, echo=FALSE, fig.width=12, fig.height=8}
capa_steps <- data.frame(
  step = 1:8,
  name = c("Identification", "Evaluation", "Investigation",
           "Root Cause", "Action Plan", "Implementation",
           "Verification", "Closure"),
  description = c(
    "Identify the nonconformity\nor potential issue",
    "Assess significance;\ncontainment if needed",
    "Gather data and\nevidence",
    "Determine root cause(s)\nusing RCA tools",
    "Define corrective/\npreventive actions",
    "Execute action plan;\nassign responsibility",
    "Verify actions completed\nand effective",
    "Document and close;\nshare lessons learned"
  ),
  y = c(4, 4, 4, 4, 2, 2, 2, 2),
  x = c(1, 2, 3, 4, 4, 3, 2, 1)
)

ggplot(capa_steps, aes(x = x, y = y)) +
  geom_tile(aes(fill = step), width = 0.9, height = 1.5, color = "white", size = 1) +
  geom_text(aes(label = paste0(step, ". ", name)), fontface = "bold",
            size = 3, color = "white", vjust = -0.5) +
  geom_text(aes(label = description), size = 2.5, color = "white",
            vjust = 1, lineheight = 0.85) +
  # Arrows showing flow
  annotate("segment", x = 1.5, xend = 1.9, y = 4, yend = 4,
           arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("segment", x = 2.5, xend = 2.9, y = 4, yend = 4,
           arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("segment", x = 3.5, xend = 3.9, y = 4, yend = 4,
           arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("segment", x = 4, xend = 4, y = 3.2, yend = 2.8,
           arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("segment", x = 3.5, xend = 3.1, y = 2, yend = 2,
           arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("segment", x = 2.5, xend = 2.1, y = 2, yend = 2,
           arrow = arrow(length = unit(0.2, "cm"))) +
  annotate("segment", x = 1.5, xend = 1.1, y = 2, yend = 2,
           arrow = arrow(length = unit(0.2, "cm"))) +
  scale_fill_gradient(low = "#27ae60", high = "#3498db", guide = "none") +
  coord_cartesian(xlim = c(0.3, 4.7), ylim = c(1, 5)) +
  labs(title = "CAPA Process Flow",
       subtitle = "Eight-step process for effective corrective and preventive action") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### CAPA Effectiveness Criteria

```{r capa-effectiveness, echo=TRUE}
# CAPA Effectiveness Tracking
set.seed(42)

# Simulate CAPA data for a year
months <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
            "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

capa_data <- data.frame(
  Month = factor(months, levels = months),
  Total_CAPAs = c(12, 15, 18, 14, 11, 13, 10, 9, 8, 7, 9, 6),
  Closed_On_Time = c(8, 10, 12, 11, 9, 11, 9, 8, 7, 6, 8, 5),
  Effective = c(7, 9, 10, 10, 8, 10, 8, 8, 7, 6, 8, 5),
  Recurrence = c(2, 1, 3, 1, 0, 1, 0, 0, 1, 0, 0, 0)
)

# Calculate metrics
capa_data$On_Time_Rate <- round(capa_data$Closed_On_Time / capa_data$Total_CAPAs * 100, 1)
capa_data$Effectiveness_Rate <- round(capa_data$Effective / capa_data$Closed_On_Time * 100, 1)

cat("CAPA Performance Metrics Summary:\n")
cat("─────────────────────────────────\n")
cat("Total CAPAs initiated:", sum(capa_data$Total_CAPAs), "\n")
cat("Closed on time:", sum(capa_data$Closed_On_Time), "(",
    round(sum(capa_data$Closed_On_Time)/sum(capa_data$Total_CAPAs)*100, 1), "%)\n")
cat("Verified effective:", sum(capa_data$Effective), "(",
    round(sum(capa_data$Effective)/sum(capa_data$Closed_On_Time)*100, 1), "%)\n")
cat("Recurrences:", sum(capa_data$Recurrence), "\n")
```

```{r capa-chart, echo=FALSE, fig.width=11, fig.height=6}
capa_long <- capa_data %>%
  select(Month, Total_CAPAs, Closed_On_Time, Effective) %>%
  pivot_longer(cols = -Month, names_to = "Metric", values_to = "Count")

ggplot(capa_long, aes(x = Month, y = Count, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_line(data = capa_data, aes(x = Month, y = Recurrence * 5, group = 1),
            color = "red", size = 1.2, inherit.aes = FALSE) +
  geom_point(data = capa_data, aes(x = Month, y = Recurrence * 5),
             color = "red", size = 3, inherit.aes = FALSE) +
  scale_fill_manual(values = c("#e74c3c", "#f39c12", "#27ae60"),
                    labels = c("Closed On Time", "Verified Effective", "Total CAPAs")) +
  scale_y_continuous(sec.axis = sec_axis(~./5, name = "Recurrences")) +
  labs(title = "CAPA Performance Tracking",
       subtitle = "Monthly CAPA metrics with recurrence trend (red line)",
       x = "", y = "Number of CAPAs", fill = "") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )
```

---

## Management Review

**Management review** is a formal evaluation of the QMS by top management to ensure its continuing suitability, adequacy, and effectiveness.

### Management Review Inputs

```{r mgmt-review-inputs, echo=FALSE}
review_inputs <- data.frame(
  Input = c("Previous Review Actions", "Changes in Issues/Context",
            "Quality Performance", "Resource Adequacy",
            "Risk/Opportunity Actions", "Improvement Opportunities"),
  Details = c(
    "Status of actions from previous management reviews",
    "Changes in external/internal issues; interested party requirements",
    "Customer satisfaction; quality objectives achievement; process performance; nonconformities; audits; supplier performance",
    "Adequacy of resources for QMS maintenance and improvement",
    "Effectiveness of actions taken to address risks and opportunities",
    "Opportunities for improvement identified from various sources"
  ),
  Data_Sources = c(
    "Previous meeting minutes; action log",
    "SWOT analysis; customer feedback; regulatory changes",
    "KPI dashboards; audit reports; CAPA data; customer complaints",
    "Budget status; staffing levels; equipment condition",
    "Risk register updates; opportunity tracking",
    "Employee suggestions; benchmarking; industry developments"
  )
)

kable(review_inputs, col.names = c("Required Input", "Details to Review", "Typical Data Sources"),
      caption = "Management Review Inputs (ISO 9001 Clause 9.3.2)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "20%")
```

### Management Review Outputs

```{r mgmt-review-outputs, echo=FALSE}
review_outputs <- data.frame(
  Output = c("Improvement Decisions", "QMS Changes", "Resource Needs"),
  Requirement = c(
    "Decisions and actions related to improvement opportunities",
    "Any need for changes to the QMS",
    "Any resource needs"
  ),
  Examples = c(
    "Launch kaizen project for top 3 customer complaints; implement new quality training program",
    "Revise control plan based on PFMEA update; modify inspection frequency",
    "Hire additional quality technician; purchase new CMM; upgrade software"
  )
)

kable(review_outputs, col.names = c("Required Output", "ISO Requirement", "Examples"),
      caption = "Management Review Outputs (ISO 9001 Clause 9.3.3)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "20%")
```

### Sample Management Review Dashboard

```{r mgmt-dashboard, echo=FALSE, fig.width=12, fig.height=10}
# Create a sample management review dashboard
par(mfrow = c(2, 2))

# Data for dashboard
set.seed(123)

# Quality Objectives
objectives <- data.frame(
  Objective = c("Customer\nComplaints", "On-Time\nDelivery", "First Pass\nYield", "Scrap\nRate"),
  Target = c(5, 98, 95, 2),
  Actual = c(3, 97.5, 96.2, 1.8),
  x = 1:4
)

p1 <- ggplot(objectives, aes(x = x)) +
  geom_bar(aes(y = Target), stat = "identity", fill = "gray80", width = 0.4, position = position_nudge(x = -0.2)) +
  geom_bar(aes(y = Actual), stat = "identity", fill = "#27ae60", width = 0.4, position = position_nudge(x = 0.2)) +
  geom_text(aes(y = 0, label = Objective), vjust = 1.5, size = 3, lineheight = 0.8) +
  labs(title = "Quality Objectives: Target vs Actual", x = "", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), plot.title = element_text(face = "bold", size = 11))

# Customer Satisfaction Trend
satisfaction <- data.frame(
  Quarter = c("Q1", "Q2", "Q3", "Q4"),
  Score = c(4.2, 4.3, 4.5, 4.6)
)

p2 <- ggplot(satisfaction, aes(x = Quarter, y = Score, group = 1)) +
  geom_line(color = "#3498db", size = 1.5) +
  geom_point(color = "#3498db", size = 4) +
  geom_hline(yintercept = 4.0, linetype = "dashed", color = "red") +
  annotate("text", x = 4.2, y = 4.0, label = "Target", color = "red", size = 3) +
  scale_y_continuous(limits = c(3.5, 5)) +
  labs(title = "Customer Satisfaction Trend", x = "", y = "Score (1-5)") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 11))

# Audit Findings
audit_data <- data.frame(
  Type = c("Major NC", "Minor NC", "Observations", "OFIs"),
  Count = c(0, 3, 5, 8)
)

p3 <- ggplot(audit_data, aes(x = Type, y = Count, fill = Type)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = Count), vjust = -0.5, size = 4) +
  scale_fill_manual(values = c("Major NC" = "#e74c3c", "Minor NC" = "#f39c12",
                               "Observations" = "#3498db", "OFIs" = "#27ae60"),
                    guide = "none") +
  labs(title = "Internal Audit Findings (YTD)", x = "", y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 11))

# CAPA Status
capa_status <- data.frame(
  Status = c("Open", "In Progress", "Verification", "Closed"),
  Count = c(5, 12, 8, 45)
)

p4 <- ggplot(capa_status, aes(x = "", y = Count, fill = Status)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(Count, "\n(", round(Count/sum(Count)*100), "%)")),
            position = position_stack(vjust = 0.5), color = "white", size = 3) +
  scale_fill_manual(values = c("#e74c3c", "#f39c12", "#3498db", "#27ae60")) +
  labs(title = "CAPA Status Distribution", fill = "") +
  theme_void() +
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5))

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2,
                        top = grid::textGrob("Management Review Dashboard",
                                             gp = grid::gpar(fontface = "bold", fontsize = 14)))
```

---

## Continuous Improvement

**Continuous improvement** is a fundamental principle of quality management - the ongoing effort to improve products, services, and processes.

### Improvement Methodologies

```{r improvement-methods, echo=FALSE}
methods <- data.frame(
  Methodology = c("PDCA", "DMAIC", "Kaizen", "A3 Problem Solving", "8D"),
  Origin = c("Deming/Shewhart", "Six Sigma", "Toyota/Lean", "Toyota", "Ford"),
  Best_For = c(
    "General improvement cycle; process changes",
    "Data-driven problem solving; reducing variation",
    "Small, incremental improvements; team-based",
    "Visual problem solving; root cause analysis",
    "Customer complaints; complex problems"
  ),
  Steps = c(
    "Plan-Do-Check-Act",
    "Define-Measure-Analyze-Improve-Control",
    "Identify waste → Improve → Standardize",
    "Background → Goals → Analysis → Countermeasures → Follow-up",
    "D0-D8 (Team, Problem, Containment, Root Cause, Actions, Verify, Prevent, Close)"
  )
)

kable(methods, caption = "Continuous Improvement Methodologies") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "15%")
```

### Key Performance Indicators (KPIs)

```{r kpi-examples, echo=FALSE}
kpis <- data.frame(
  Category = c(rep("Quality", 4), rep("Delivery", 3), rep("Cost", 3), rep("Safety", 2)),
  KPI = c("First Pass Yield (FPY)", "Customer Complaints (PPM)", "Internal Reject Rate", "Cost of Quality",
          "On-Time Delivery (OTD)", "Lead Time", "Order Fulfillment Rate",
          "Scrap Rate", "Rework Cost", "Warranty Cost",
          "Recordable Incident Rate", "Near Miss Reports"),
  Formula = c(
    "(Good units first time / Total units) × 100%",
    "(Complaints / Units shipped) × 1,000,000",
    "(Rejected units / Total inspected) × 100%",
    "Prevention + Appraisal + Internal Failure + External Failure",
    "(Orders delivered on time / Total orders) × 100%",
    "Order date to delivery date",
    "(Complete orders shipped / Total orders) × 100%",
    "(Scrap value / Total production value) × 100%",
    "Labor + Material for rework",
    "Warranty claims cost / Revenue",
    "(Recordable incidents × 200,000) / Hours worked",
    "Count of near miss reports"
  ),
  Target = c("≥98%", "<100", "<1%", "<3% of sales",
             "≥98%", "Industry benchmark", "≥99%",
             "<2%", "Minimize", "<0.5% of sales",
             "<1.0", "Encourage reporting")
)

kable(kpis, caption = "Common Quality Management KPIs") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "10%") %>%
  pack_rows("Quality Metrics", 1, 4, background = "#ebf5fb") %>%
  pack_rows("Delivery Metrics", 5, 7, background = "#eafaf1") %>%
  pack_rows("Cost Metrics", 8, 10, background = "#fef9e7") %>%
  pack_rows("Safety Metrics", 11, 12, background = "#fdedec")
```

### Cost of Quality (COQ)

```{r coq-analysis, echo=TRUE}
# Cost of Quality Analysis
coq_data <- data.frame(
  Category = c("Prevention", "Appraisal", "Internal Failure", "External Failure"),
  Description = c("Quality planning, training, process control",
                  "Inspection, testing, audits, calibration",
                  "Scrap, rework, reinspection, downtime",
                  "Warranty, returns, complaints, recalls"),
  Cost = c(50000, 75000, 120000, 180000)
)

total_coq <- sum(coq_data$Cost)
revenue <- 10000000  # $10M revenue

coq_data$Percent_of_COQ <- round(coq_data$Cost / total_coq * 100, 1)
coq_data$Percent_of_Revenue <- round(coq_data$Cost / revenue * 100, 2)

cat("Cost of Quality Analysis:\n")
cat("─────────────────────────────\n")
print(coq_data[, c("Category", "Cost", "Percent_of_COQ")])
cat("─────────────────────────────\n")
cat("Total Cost of Quality: $", format(total_coq, big.mark = ","), "\n")
cat("COQ as % of Revenue:", round(total_coq/revenue*100, 1), "%\n\n")

# Analysis
conformance <- coq_data$Cost[1] + coq_data$Cost[2]
nonconformance <- coq_data$Cost[3] + coq_data$Cost[4]

cat("Conformance Costs (Prevention + Appraisal): $", format(conformance, big.mark = ","),
    "(", round(conformance/total_coq*100, 1), "%)\n")
cat("Nonconformance Costs (Failures): $", format(nonconformance, big.mark = ","),
    "(", round(nonconformance/total_coq*100, 1), "%)\n")
cat("\nRecommendation: Increase prevention spending to reduce failure costs\n")
```

```{r coq-chart, echo=FALSE, fig.width=10, fig.height=6}
coq_data$Category <- factor(coq_data$Category,
                            levels = c("Prevention", "Appraisal", "Internal Failure", "External Failure"))

ggplot(coq_data, aes(x = Category, y = Cost/1000, fill = Category)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_text(aes(label = paste0("$", Cost/1000, "K\n(", Percent_of_COQ, "%)")),
            vjust = -0.2, size = 3.5) +
  scale_fill_manual(values = c("#27ae60", "#3498db", "#f39c12", "#e74c3c"), guide = "none") +
  scale_y_continuous(limits = c(0, 220), labels = scales::dollar_format(suffix = "K")) +
  annotate("rect", xmin = 0.5, xmax = 2.5, ymin = -10, ymax = -5, fill = "#27ae60", alpha = 0.3) +
  annotate("text", x = 1.5, y = -7.5, label = "Conformance Costs", size = 3) +
  annotate("rect", xmin = 2.5, xmax = 4.5, ymin = -10, ymax = -5, fill = "#e74c3c", alpha = 0.3) +
  annotate("text", x = 3.5, y = -7.5, label = "Nonconformance Costs", size = 3) +
  labs(title = "Cost of Quality Analysis",
       subtitle = paste0("Total COQ: $", format(total_coq, big.mark = ","),
                        " (", round(total_coq/revenue*100, 1), "% of revenue)"),
       x = "", y = "Cost") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(size = 11)
  )
```

---

## Integration with Quality Tools

A mature QMS integrates various quality tools throughout the organization.

### Quality Tools Integration Map

```{r tools-integration, echo=FALSE, fig.width=14, fig.height=9}
tools_map <- data.frame(
  Phase = c("Design", "Design", "Launch", "Launch", "Production", "Production", "Production", "Improvement"),
  Tool = c("DFMEA", "DFM/DFA", "PFMEA", "Control Plan", "SPC", "MSA", "Audit", "CAPA"),
  QMS_Link = c("Risk assessment (6.1)", "Design controls (8.3)", "Risk assessment (6.1)",
               "Operational control (8.1)", "Monitoring (9.1)", "Resources (7.1.5)",
               "Internal audit (9.2)", "Nonconformity (10.2)"),
  x = c(1, 1.5, 2.5, 3, 4, 4.5, 5.5, 6.5),
  y = c(2, 1, 2, 1, 2, 1, 2, 1.5)
)

ggplot(tools_map, aes(x = x, y = y)) +
  # Phase boxes
  annotate("rect", xmin = 0.5, xmax = 1.8, ymin = 0.3, ymax = 2.7,
           fill = "#3498db", alpha = 0.2, color = "#3498db") +
  annotate("text", x = 1.15, y = 2.5, label = "DESIGN", fontface = "bold", color = "#3498db") +

  annotate("rect", xmin = 2, xmax = 3.5, ymin = 0.3, ymax = 2.7,
           fill = "#27ae60", alpha = 0.2, color = "#27ae60") +
  annotate("text", x = 2.75, y = 2.5, label = "LAUNCH", fontface = "bold", color = "#27ae60") +

  annotate("rect", xmin = 3.7, xmax = 6, ymin = 0.3, ymax = 2.7,
           fill = "#f39c12", alpha = 0.2, color = "#f39c12") +
  annotate("text", x = 4.85, y = 2.5, label = "PRODUCTION", fontface = "bold", color = "#f39c12") +

  annotate("rect", xmin = 6.2, xmax = 7, ymin = 0.3, ymax = 2.7,
           fill = "#9b59b6", alpha = 0.2, color = "#9b59b6") +
  annotate("text", x = 6.6, y = 2.5, label = "IMPROVE", fontface = "bold", color = "#9b59b6") +

  # Tool boxes
  geom_label(aes(label = Tool), fill = "white", fontface = "bold", size = 3.5) +
  geom_text(aes(label = QMS_Link, y = y - 0.35), size = 2.5, color = "gray40") +

  # Flow arrows
  annotate("segment", x = 1.8, xend = 2.3, y = 1.5, yend = 1.5,
           arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +
  annotate("segment", x = 3.3, xend = 3.8, y = 1.5, yend = 1.5,
           arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +
  annotate("segment", x = 5.8, xend = 6.2, y = 1.5, yend = 1.5,
           arrow = arrow(length = unit(0.2, "cm")), color = "gray50") +

  # Feedback loop
  annotate("curve", x = 6.6, xend = 1.15, y = 0.5, yend = 0.5,
           curvature = 0.3, arrow = arrow(length = unit(0.2, "cm")),
           color = "#e74c3c", linetype = "dashed") +
  annotate("text", x = 4, y = 0.15, label = "Lessons Learned / Continuous Improvement",
           color = "#e74c3c", fontface = "italic", size = 3) +

  coord_cartesian(xlim = c(0, 7.5), ylim = c(-0.2, 3)) +
  labs(title = "Quality Tools Integration Throughout Product Lifecycle",
       subtitle = "Each tool connects to specific ISO 9001 requirements") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### Linking Tools to QMS Requirements

```{r tools-qms-link, echo=FALSE}
tool_links <- data.frame(
  Tool = c("PFMEA", "Control Plan", "SPC", "MSA", "Internal Audit", "CAPA", "Management Review"),
  ISO_Clause = c("6.1, 8.1", "8.1, 8.5.1", "9.1.1", "7.1.5", "9.2", "10.2", "9.3"),
  Purpose = c(
    "Identify and mitigate process risks before production",
    "Define controls for special characteristics",
    "Monitor process performance; detect special causes",
    "Ensure measurement system is adequate",
    "Verify QMS conformance and effectiveness",
    "Eliminate causes of nonconformities",
    "Evaluate QMS suitability and effectiveness"
  ),
  Output_Used_By = c(
    "Control Plan, Work Instructions",
    "Work Instructions, SPC requirements",
    "CAPA triggers, Management Review",
    "SPC validity, Control Plan",
    "CAPA triggers, Management Review",
    "Process improvements, Lessons learned",
    "Strategic decisions, Resource allocation"
  )
)

kable(tool_links, col.names = c("Quality Tool", "ISO 9001 Clause", "Purpose", "Output Feeds Into"),
      caption = "Quality Tools and QMS Integration") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "15%")
```

---

## Certification Process

Organizations can seek **third-party certification** to demonstrate QMS conformance to customers and stakeholders.

### The Certification Journey

```{r certification-process, echo=FALSE, fig.width=12, fig.height=6}
cert_steps <- data.frame(
  step = 1:8,
  name = c("Gap Analysis", "Implementation", "Internal Audit",
           "Management Review", "Registrar Selection", "Stage 1 Audit",
           "Stage 2 Audit", "Certification"),
  duration = c("1-2 months", "3-9 months", "1 month", "Ongoing",
               "1 month", "1 day", "2-4 days", "3 years"),
  x = 1:8
)

ggplot(cert_steps, aes(x = x, y = 1)) +
  geom_segment(aes(x = 1, xend = 8, y = 1, yend = 1), color = "gray70", size = 2) +
  geom_point(size = 12, color = "#3498db") +
  geom_text(aes(label = step), color = "white", fontface = "bold", size = 4) +
  geom_text(aes(label = name, y = 1.4), size = 3, fontface = "bold", angle = 45, hjust = 0) +
  geom_text(aes(label = duration, y = 0.6), size = 2.5, color = "gray50") +
  scale_y_continuous(limits = c(0.3, 2)) +
  labs(title = "Path to ISO 9001 Certification",
       subtitle = "Typical timeline: 6-18 months depending on organization readiness") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )
```

### Certification Audit Types

```{r audit-types, echo=FALSE}
audit_types <- data.frame(
  Audit = c("Stage 1 (Documentation)", "Stage 2 (Certification)",
            "Surveillance", "Recertification"),
  Purpose = c(
    "Review QMS documentation; verify readiness for Stage 2",
    "Evaluate implementation and effectiveness of QMS",
    "Verify continued conformance (annual)",
    "Full re-audit at end of 3-year cycle"
  ),
  Focus = c(
    "Quality manual, procedures, scope, internal audits, management review",
    "All clauses; process effectiveness; records; interviews",
    "Selected processes; previous findings; changes",
    "Complete review similar to Stage 2"
  ),
  Typical_Duration = c(
    "0.5-1 day (remote or on-site)",
    "2-5 days depending on size",
    "1-2 days annually",
    "2-4 days"
  )
)

kable(audit_types, col.names = c("Audit Type", "Purpose", "Focus Areas", "Duration"),
      caption = "Third-Party Certification Audit Types") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "18%")
```

---

## Video Resources

### ISO 9001:2015 Overview

<iframe width="560" height="315" src="https://www.youtube.com/embed/O5T4H8K_rwQ" title="ISO 9001:2015 Overview" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

### Internal Auditing Best Practices

<iframe width="560" height="315" src="https://www.youtube.com/embed/deRqslBeMrE" title="Internal Auditing" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

---

## Summary

A Quality Management System provides the framework for consistently meeting customer requirements and improving organizational performance:

1. **ISO 9001:2015** provides the foundational requirements using risk-based thinking and PDCA
2. **Industry standards** (IATF 16949, AS9100, ISO 22000) add sector-specific requirements
3. **Documentation** establishes the hierarchy from policy to procedures to records
4. **Internal audits** verify conformance and identify improvement opportunities
5. **CAPA** eliminates root causes of nonconformities and prevents recurrence
6. **Management review** ensures QMS suitability and drives strategic improvement
7. **Continuous improvement** is embedded through KPIs, COQ analysis, and improvement methodologies
8. **Quality tools** (PFMEA, SPC, MSA) integrate throughout the product lifecycle

> "Quality is not an act, it is a habit." — Aristotle

---

## Review Questions

<details><summary>**Question 1**: Explain the seven quality management principles and how they relate to each other.</summary>

**Answer:**

The seven quality management principles form an interconnected foundation:

1. **Customer Focus** (Central principle)
   - Understanding and meeting customer needs is the primary purpose
   - All other principles support this goal

2. **Leadership**
   - Leaders establish unity of purpose and direction
   - Create conditions for people to achieve quality objectives
   - Enables all other principles

3. **Engagement of People**
   - Competent, empowered people at all levels
   - Essential for effective processes and customer focus

4. **Process Approach**
   - Activities managed as interrelated processes
   - Enables consistent, predictable results

5. **Improvement**
   - Ongoing focus on enhancing performance
   - Essential for maintaining customer satisfaction

6. **Evidence-based Decision Making**
   - Decisions based on analysis of data
   - Reduces uncertainty; improves decision quality

7. **Relationship Management**
   - Managing relationships with interested parties (suppliers, partners)
   - Optimizes their impact on performance

**Relationships:**
- Customer Focus drives the need for all others
- Leadership enables and supports all principles
- Engagement of People is essential for Process Approach and Improvement
- Evidence-based Decision Making supports Improvement
- Relationship Management extends quality beyond the organization
</details>

<details><summary>**Question 2**: Compare ISO 9001, IATF 16949, and AS9100. When would each be required?</summary>

**Answer:**

| Standard | Base | Industry | Additional Focus |
|----------|------|----------|------------------|
| **ISO 9001:2015** | Standalone | General | Quality fundamentals |
| **IATF 16949:2016** | ISO 9001 + automotive | Automotive | PPAP, APQP, Core Tools |
| **AS9100D** | ISO 9001 + aerospace | Aerospace/Defense | Configuration mgmt, FAI, counterfeit prevention |

**When Required:**

**ISO 9001:**
- Any organization wanting to demonstrate quality capability
- Customer or contract requirement for QMS certification
- Basis for all other standards
- Suitable for service, manufacturing, any industry

**IATF 16949:**
- Suppliers to automotive OEMs (Ford, GM, Toyota, VW, etc.)
- Required for production parts/materials
- Required throughout automotive supply chain
- Includes Customer-Specific Requirements (CSRs)

**AS9100:**
- Suppliers to aerospace/defense contractors
- Boeing, Airbus, Lockheed Martin, military programs
- Required for flight-critical components
- OASIS database registration required

**Key Differences:**
- IATF 16949: Emphasizes PPAP, APQP, manufacturing process control
- AS9100: Emphasizes traceability, configuration control, first article inspection
- Both build on ISO 9001 with sector-specific additions
</details>

<details><summary>**Question 3**: Describe the documentation hierarchy in a QMS. Give an example of each level.</summary>

**Answer:**

**Level 1: Quality Policy and Objectives**
- Strategic direction for quality
- Signed by top management
- Communicated to all employees
- *Example:* "ABC Company is committed to delivering defect-free products that meet customer requirements through continuous improvement of our processes and people."

**Level 2: Quality Manual and Procedures**
- Describes WHAT is done and WHO is responsible
- System-level documents
- *Example:* Document Control Procedure (QP-001)
  - Purpose: Control creation, approval, distribution of documents
  - Scope: All QMS documents
  - Responsibilities: Document Control Coordinator, Department Managers
  - Process flow for document changes

**Level 3: Work Instructions and Specifications**
- Describes HOW tasks are performed
- Detailed step-by-step instructions
- *Example:* Work Instruction for CNC Lathe Setup (WI-MAC-015)
  - Step 1: Verify program number matches work order
  - Step 2: Install correct tooling per setup sheet
  - Step 3: Set tool offsets using touch-off procedure
  - Step 4: Run first piece and verify dimensions

**Level 4: Records and Forms**
- Evidence that activities were performed
- Completed forms, inspection results, logs
- *Example:* First Article Inspection Report (FM-QC-008)
  - Part number, revision, date
  - Dimensional measurements vs. specifications
  - Inspector signature
  - Disposition (Accept/Reject)
</details>

<details><summary>**Question 4**: What are the key elements of an effective internal audit program?</summary>

**Answer:**

**1. Audit Program Planning**
- Annual audit schedule covering all QMS processes
- Risk-based frequency (higher risk = more frequent)
- Documented audit program procedure
- Resource allocation (trained auditors, time)

**2. Auditor Competence**
- Trained in audit techniques
- Knowledge of standards (ISO 9001, industry-specific)
- Understanding of processes being audited
- Independence from area being audited

**3. Audit Preparation**
- Review of relevant documents (procedures, previous audits)
- Prepared checklist linked to requirements
- Notification to auditee
- Defined scope and objectives

**4. Audit Execution**
- Opening meeting (purpose, scope, schedule)
- Evidence collection (documents, records, interviews, observation)
- Objective evaluation against criteria
- Note taking and evidence documentation

**5. Reporting**
- Clear finding statements (nonconformities, observations)
- Reference to requirement violated
- Objective evidence cited
- Closing meeting to present findings

**6. Follow-up**
- Corrective action requests issued
- Root cause analysis required
- Verification of corrective action implementation
- Verification of effectiveness

**7. Continuous Improvement of Audit Program**
- Evaluate audit program effectiveness
- Update based on organizational changes
- Calibrate auditors periodically
- Incorporate lessons learned
</details>

<details><summary>**Question 5**: Explain the difference between corrective action and preventive action. Provide an example of each.</summary>

**Answer:**

**Corrective Action:**
- Eliminates the cause of a *detected* nonconformity
- Reactive - responds to something that has already happened
- Goal: Prevent recurrence of the specific problem

*Example:*
- **Nonconformity:** Customer received shipment with wrong part numbers (3 incidents in past month)
- **Root Cause:** Shipping labels printed from outdated part number list
- **Corrective Action:**
  1. Update master part number list
  2. Link shipping system directly to ERP for current data
  3. Add verification step comparing label to packing list
- **Verification:** No recurrence after 3 months

**Preventive Action:**
- Eliminates the cause of a *potential* nonconformity
- Proactive - addresses something that hasn't happened yet
- Goal: Prevent occurrence before it happens

*Example:*
- **Potential Nonconformity:** New CNC machine may produce out-of-spec parts due to operator unfamiliarity
- **Risk Identified Through:** PFMEA analysis during new equipment installation
- **Preventive Action:**
  1. Develop comprehensive training program before machine goes live
  2. Create detailed setup procedures with photos
  3. Implement first-article inspection for first 30 days
  4. Assign experienced mentor to new operators
- **Verification:** Track first-pass yield during initial production period

**Key Difference:**
- Corrective = Fix what went wrong
- Preventive = Stop what might go wrong

**Note:** ISO 9001:2015 doesn't explicitly require "preventive action" as a separate process. Instead, risk-based thinking throughout the QMS serves the preventive function.
</details>

<details><summary>**Question 6**: Calculate and analyze the Cost of Quality for the following data. What recommendations would you make?</summary>

| Category | Cost |
|----------|------|
| Quality planning | $35,000 |
| Training | $25,000 |
| Incoming inspection | $40,000 |
| In-process inspection | $55,000 |
| Final inspection | $30,000 |
| Scrap | $85,000 |
| Rework | $65,000 |
| Warranty claims | $120,000 |
| Customer returns | $45,000 |

Annual revenue: $8,000,000

**Answer:**

```{r q6-answer, echo=TRUE}
# Cost of Quality Analysis
prevention <- 35000 + 25000  # Planning + Training
appraisal <- 40000 + 55000 + 30000  # Inspections
internal_failure <- 85000 + 65000  # Scrap + Rework
external_failure <- 120000 + 45000  # Warranty + Returns

total_coq <- prevention + appraisal + internal_failure + external_failure
revenue <- 8000000

cat("Cost of Quality Breakdown:\n")
cat("─────────────────────────────\n")
cat("Prevention:       $", format(prevention, big.mark = ","),
    "(", round(prevention/total_coq*100, 1), "%)\n")
cat("Appraisal:        $", format(appraisal, big.mark = ","),
    "(", round(appraisal/total_coq*100, 1), "%)\n")
cat("Internal Failure: $", format(internal_failure, big.mark = ","),
    "(", round(internal_failure/total_coq*100, 1), "%)\n")
cat("External Failure: $", format(external_failure, big.mark = ","),
    "(", round(external_failure/total_coq*100, 1), "%)\n")
cat("─────────────────────────────\n")
cat("Total COQ:        $", format(total_coq, big.mark = ","), "\n")
cat("COQ % of Revenue:", round(total_coq/revenue*100, 1), "%\n\n")

conformance <- prevention + appraisal
nonconformance <- internal_failure + external_failure

cat("Conformance Costs:    $", format(conformance, big.mark = ","),
    "(", round(conformance/total_coq*100, 1), "%)\n")
cat("Nonconformance Costs: $", format(nonconformance, big.mark = ","),
    "(", round(nonconformance/total_coq*100, 1), "%)\n")
```

**Analysis:**
- COQ at 7.5% of revenue is HIGH (target <3%)
- External failure costs are the largest category (27.5%)
- Nonconformance costs (60.8%) greatly exceed conformance (39.2%)
- Prevention is only 10% of COQ - very low

**Recommendations:**
1. **Increase Prevention Investment:**
   - Implement SPC on critical processes to reduce defects
   - Enhance operator training on quality standards
   - Apply PFMEA to identify and control failure modes

2. **Reduce External Failures:**
   - Analyze warranty data for root causes
   - Improve final inspection effectiveness
   - Implement containment for known issues

3. **Optimize Appraisal:**
   - Use SPC to reduce inspection where processes are capable
   - Implement risk-based inspection (more for critical, less for proven)

4. **Target COQ Reduction:**
   - Short-term: Reduce external failures by 50% ($82,500 savings)
   - Medium-term: Reduce internal failures by 30% ($45,000 savings)
   - Long-term: Achieve COQ <3% of revenue ($240,000)
</details>

<details><summary>**Question 7**: What is risk-based thinking in ISO 9001:2015? How is it different from formal risk management?</summary>

**Answer:**

**Risk-Based Thinking in ISO 9001:2015:**

Risk-based thinking is a way of considering risks and opportunities when designing, implementing, and operating the QMS. It:

- Applies throughout the standard (not a separate clause)
- Is proportional to potential impact on conformity
- Makes preventive action part of routine planning
- Considers both risks (negative) and opportunities (positive)

**Where Risk-Based Thinking Applies:**

| Clause | Application |
|--------|-------------|
| 4.1 Context | Identify external/internal issues affecting QMS |
| 4.2 Interested Parties | Understand requirements that could impact quality |
| 6.1 Planning | Plan actions to address risks and opportunities |
| 8.1 Operations | Consider risks in operational planning |
| 9.1 Performance | Monitor effectiveness of risk actions |
| 10.2 Corrective Action | Update risks/opportunities based on nonconformities |

**How It Differs from Formal Risk Management:**

| Aspect | Risk-Based Thinking | Formal Risk Management |
|--------|---------------------|------------------------|
| Requirement | Implicit throughout ISO 9001 | Explicit (ISO 31000) |
| Documentation | No formal risk register required | Documented risk register |
| Methodology | No specific method required | Defined methodology (FMEA, etc.) |
| Quantification | Not required | Often quantitative |
| Scope | QMS processes | Enterprise-wide |

**Practical Implementation:**

Organizations should:
1. Consider risks when designing processes
2. Build controls proportional to risk level
3. Monitor and review effectiveness
4. Update as context changes

ISO 9001 does NOT require:
- Formal risk management process
- Risk register
- Specific risk assessment methodology
- Documented risk assessments (unless organization decides to)

However, many organizations do use tools like PFMEA to fulfill risk-based thinking requirements, especially in automotive (IATF 16949 requires it).
</details>

<details><summary>**Question 8**: Describe what management review is and what should be included in the inputs and outputs.</summary>

**Answer:**

**What is Management Review?**

Management review is a formal, periodic evaluation of the QMS by top management to:
- Ensure continuing suitability (appropriate for the organization)
- Ensure adequacy (sufficient to meet requirements)
- Ensure effectiveness (achieving intended results)
- Ensure alignment with strategic direction
- Identify improvement opportunities

**Frequency:** At least annually; many organizations conduct quarterly reviews

**Required Inputs (ISO 9001 Clause 9.3.2):**

| Input | What to Review |
|-------|----------------|
| Status of previous actions | Actions from prior reviews - completed? effective? |
| Changes in external/internal issues | Market changes, regulations, technology, organizational changes |
| QMS performance and effectiveness | Customer satisfaction, quality objectives, process performance, nonconformities, audit results, supplier performance |
| Resource adequacy | People, infrastructure, environment, knowledge |
| Effectiveness of risk actions | Did actions address risks/opportunities? |
| Improvement opportunities | Suggestions, benchmarking, new technologies |

**Required Outputs (ISO 9001 Clause 9.3.3):**

| Output | Examples |
|--------|----------|
| Improvement decisions | Launch specific improvement projects, address recurring issues |
| QMS changes | Revise quality objectives, update processes, modify documentation |
| Resource needs | Budget approvals, hiring, equipment purchases, training |

**Best Practices:**
- Use a standard agenda aligned with required inputs
- Prepare data in advance (dashboards, trend charts)
- Document decisions and action items with owners and due dates
- Follow up on action completion
- Keep minutes as quality records

**Evidence Required:**
- Meeting minutes or management review report
- Attendance record (top management participation)
- Action items with status tracking
- Data/reports presented
</details>

---

## References

1. ISO 9001:2015. *Quality Management Systems - Requirements*. International Organization for Standardization.

2. ISO 9000:2015. *Quality Management Systems - Fundamentals and Vocabulary*. International Organization for Standardization.

3. IATF 16949:2016. *Quality Management System Requirements for Automotive Production and Relevant Service Parts Organizations*. International Automotive Task Force.

4. AS9100D:2016. *Quality Management Systems - Requirements for Aviation, Space and Defense Organizations*. SAE International.

5. ISO 22000:2018. *Food Safety Management Systems - Requirements for Any Organization in the Food Chain*. International Organization for Standardization.

6. ISO 19011:2018. *Guidelines for Auditing Management Systems*. International Organization for Standardization.

7. Hoyle, D. (2017). *ISO 9000 Quality Systems Handbook* (7th ed.). Routledge.

8. AIAG. (2010). *Advanced Product Quality Planning and Control Plan (APQP)* (2nd ed.). Automotive Industry Action Group.

9. ASQ. (2019). *The Certified Quality Auditor Handbook* (4th ed.). ASQ Quality Press.

10. Juran, J.M., & De Feo, J.A. (2010). *Juran's Quality Handbook* (6th ed.). McGraw-Hill.
